### 2021-03-09 ST563-FP.Rmd
initial commit

---
title: "ST563 Final Project"
author: "Mana Azizsoltani"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  pdf_document:
    toc: true
    toc_depth: 1
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r cars}
summary(cars)
```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

### 2021-03-09 ST563-FP.Rmd
knocking some basics out

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+# Set default chunk options
+knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
+
+# Load necessary libraries
+library(tidyverse) # so many things
+library(knitr) # making nice tables
+library(caret) # training models & CV
+library(leaps) # subset selection
+library(corrplot) # visualize correlation
+library(rmarkdown) # for output documents
+library(rpart) # fitting classification tree
+library(rpart.plot) # visualization of classification tree
+library(randomForest) # fitting random forest model
+library(kernlab) # fitting SVM
+library(class) # fitting kNN model
+
+# Read in data set
+wine <- read.csv2(file = "winequality-red.csv", header = TRUE)
 ```
 
-## R Markdown
+## Required Libraries
+To run the code for the project, the following libraries are required:
 
-This is an R Markdown document. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
+  * `caret`: to do the heavy lifting of training and tuning the models
+  * `tidyverse`: for all the data reading and wrangling
+  * `knitr`: for rmarkdown table outputs
+  * `rmarkdown`: for output documents
+  * `leaps`: for subset selection
+  * `rpart`: fitting the basic classification tree
+  * `rpart.plot`: visualization of the classification tree
+  * `randomForest`: fitting random forest models
+  * `kernlab`: fitting the support vector machine
+  * `class`: fitting the k-nearest neighbor model
 
-When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.
-You can embed an R code chunk like this:
+\newpage
 
-```{r cars}
-summary(cars)
-```
+## Variable Descriptions
 
-## Including Plots
+# Methods  
+## Data Processing  
+Fortunately, the `wine` data set that we worked with was very tidy; there was no missing values.
+We didn't standardize the data initially, but when training the support vector machine as well as the k-nearest neighbors models we centered and scaled the data.
 
-You can also embed plots, for example:
+As mentioned in the introduction, our focus is on the classification accuracy of machine learning methods.
+In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the five models.
+We split the data into a training and test data set in order to later evaluate the model's prediction accuracy.
+For this project we used a 75/25 split, training the data on the 75% and testing the trained models on the withheld 25%.
+We will then repeat this process over 5 folds of the data, averaging the results.
 
-```{r pressure, echo=FALSE}
-plot(pressure)
-```
+The tree-based and the K-nearest neighbors models require parameters to be tuned (more on those later).
+Since we used the `caret` package in R to fit all of our models, we used the tunes of the parameters that we used were deemed the "best" by the `train()` function.
 
-Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
+## Variable Selection  
+After running an initial random forest model on the entire dataset with selected predictors, we looked at the variable importance of the predictors to decide which variables we would use in our final model.
+Variable importance, in the context on machine learning, refers to how much a given model "uses" that variable to make accurate predictions.
+In other words, the more a model relies on a variable to make predictions, the more important it is for the model.

### 2021-03-09 ST563-FP.Rmd
added small introduction read in the data properly set data up for classification partition data

@@
 # Set default chunk options
-knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
 
@@
 # Read in data set
-wine <- read.csv2(file = "winequality-red.csv", header = TRUE)
+wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
+                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))
 ```
+
+## Introduction
+There has been a push towards the implementation of different analytics solutions within the agriculture industry, especially in the realms of crop health and yield.
+These sort of analyses can help farmers to reduce waste and improve profits.
+In particular, the viniculture industry stands to benefit from such methods, being that the production of wine is multifaceted.
+Winemakers can work to optimize revenue by analyzing their input and output (crops and wines).
+Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality.
+
+In this study, our objective is to try and predict the quality of wine using a variety of statistical learning techniques.
+We will look at this problem from a regression standpoint as well as a classification standpoint.
+Using regression, we will try and predict the exact rating (0 to 10) of a particular wine.
+On the other hand, when we look at classification, we will try to classify a particular wine as either good or bad, taking wines rated from 0-5 as low quality wines and wines rated from 6-10 as high quality wines.
 
@@
   * `rmarkdown`: for output documents
+  * `corrplot`: for correlation structure visualization
   * `leaps`: for subset selection
@@
 # Methods  
-## Data Processing  
+## Data Processing & Partitioning 
+```{r partition}
+# Prepare data for classification
+wine$class <- ifelse(wine$quality <= 5, "low", "high") %>% as.factor()
+
+# Create Data Partition
+ind <- createDataPartition(wine$quality, p = .75, list = FALSE)
+
+# Create train/test data for H1
+trn <- wine[ind,]
+tst <- wine[-ind,]
+```
+
 Fortunately, the `wine` data set that we worked with was very tidy; there was no missing values.

### 2021-03-10 ST563-FP.Rmd
added variable descriptions added table of methods

@@
 
-## Introduction
+# Introduction
 There has been a push towards the implementation of different analytics solutions within the agriculture industry, especially in the realms of crop health and yield.
 These sort of analyses can help farmers to reduce waste and improve profits.
-In particular, the viniculture industry stands to benefit from such methods, being that the production of wine is multifaceted.
+In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is multifaceted.
 Winemakers can work to optimize revenue by analyzing their input and output (crops and wines).
@@
 On the other hand, when we look at classification, we will try to classify a particular wine as either good or bad, taking wines rated from 0-5 as low quality wines and wines rated from 6-10 as high quality wines.
+We will be assessing the prediction accuracy of the following methods:
 
-## Required Libraries
+```{r methods}
+cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
+        "KNN Classification", "Logistic Regression")
+reg <- c("Regression Tree", "Boosted Tree", "Multiple Linear Regression (MLR)", 
+          "KNN Regression", "LASSO Regression")
+kable(data.frame("Classification" = cl, "Regression" = reg))
+```
+
+# Required Libraries
 To run the code for the project, the following libraries are required:
@@
 \newpage
+# Data
+The data set used is the `Wine Quality` dataset from the UCI Machine Learning Repository.
+The data set is composed of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine.
+For each wine, various features of the wine were measured, including a measurement of quality, given by an expert wine taster.
+Our objective it to use the different features to predict the quality of the wine.
 
 ## Variable Descriptions  
+Each entry in the dataset represents the different metrics of the following attributes of a single type of wine.
+
+  * fixed acidity: the quantity of fixed acids found in the wines.
+The predominant fixed acids found in wines are tartaric, malic, citric, and succinic, all of which come from the grapes with the exception of the succinic acid, which comes from the yeast in fermentation.
+  * volatile acidity: the steam distillable acids present in wine, primarily acetic acid but also lactic, formic, butyric, and propionic acids.
+These acids generally come from the fermentation process.
+  * citric acid: added to the wine as a natural preservative or for acidity and tartness.
+  * residual sugar: the quantity of sugar left in the wine after the fermentation process.
+  * chlorides: the amount of salt in a wine.
+  * free sulfur dioxide: the amount of $SO_2$ that is not bound to other molecules. 
+  * total sulfur dioxide: total amount of $SO_2$ in the wine. Sulfur Dioxide is used throughout all stages of the winemaking process to prevent oxidation and bacteria growth.
+  * density: density of the wine
+  * pH: pH of the wine
+  * sulphates: quantity of sulphates in the wine. Sulphates are used as a preservative.
+  * alcohol: alcohol content by volume.
+  * quality: score of quality of the wine given by expert tasters (score between 0 and 10).
 
 # Methods  
+## Data
 ## Data Processing & Partitioning 
@@
 As mentioned in the introduction, our focus is on the classification accuracy of machine learning methods.
-In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the five models.
+In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the models.
 We split the data into a training and test data set in order to later evaluate the model's prediction accuracy.

### 2021-03-10 ST563-FP.Rmd
Added formatting to variable list and edited methods section slightly

@@
 
+\newpage
+
 # Introduction
@@
           "KNN Regression", "LASSO Regression")
-kable(data.frame("Classification" = cl, "Regression" = reg))
+kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods")
 ```
@@
 \newpage
+
 # Data
@@
 
-  * fixed acidity: the quantity of fixed acids found in the wines.
+  * **fixed acidity**: the quantity of fixed acids found in the wines.
 The predominant fixed acids found in wines are tartaric, malic, citric, and succinic, all of which come from the grapes with the exception of the succinic acid, which comes from the yeast in fermentation.
-  * volatile acidity: the steam distillable acids present in wine, primarily acetic acid but also lactic, formic, butyric, and propionic acids.
+  * **volatile acidity**: the steam distillable acids present in wine, primarily acetic acid but also lactic, formic, butyric, and propionic acids.
 These acids generally come from the fermentation process.
-  * citric acid: added to the wine as a natural preservative or for acidity and tartness.
-  * residual sugar: the quantity of sugar left in the wine after the fermentation process.
-  * chlorides: the amount of salt in a wine.
-  * free sulfur dioxide: the amount of $SO_2$ that is not bound to other molecules. 
-  * total sulfur dioxide: total amount of $SO_2$ in the wine. Sulfur Dioxide is used throughout all stages of the winemaking process to prevent oxidation and bacteria growth.
-  * density: density of the wine
-  * pH: pH of the wine
-  * sulphates: quantity of sulphates in the wine. Sulphates are used as a preservative.
-  * alcohol: alcohol content by volume.
-  * quality: score of quality of the wine given by expert tasters (score between 0 and 10).
+  * **citric acid**: added to the wine as a natural preservative or for acidity and tartness.
+  * **residual sugar**: the quantity of sugar left in the wine after the fermentation process.
+  * **chlorides**: the amount of salt in a wine.
+  * **free sulfur dioxide**: the amount of $SO_2$ that is not bound to other molecules. 
+  * **total sulfur dioxide**: total amount of $SO_2$ in the wine. Sulfur Dioxide is used throughout all stages of the winemaking process to prevent oxidation and bacteria growth.
+  * **density**: density of the wine.
+  * **pH**: pH of the wine.
+  * **sulphates**: quantity of sulphates in the wine. Sulphates are used as a preservative.
+  * **alcohol**: alcohol content by volume.
+  * **quality**: score of quality of the wine given by expert tasters (score between 0 and 10).
 
 # Methods  
-## Data
 ## Data Processing & Partitioning 
@@
 
-The tree-based and the K-nearest neighbors models require parameters to be tuned (more on those later).
-Since we used the `caret` package in R to fit all of our models, we used the tunes of the parameters that we used were deemed the "best" by the `train()` function.
+The tree-based, LASSO, and K-nearest neighbors models require parameters to be tuned (more on those later).
+Since we used the `caret` package in R to fit all of our models, we used the tunes of the parameters that were deemed "best" by the `train()` function.
 
 ## Variable Selection
-After running an initial random forest model on the entire dataset with selected predictors, we looked at the variable importance of the predictors to decide which variables we would use in our final model.
-Variable importance, in the context on machine learning, refers to how much a given model "uses" that variable to make accurate predictions.
-In other words, the more a model relies on a variable to make predictions, the more important it is for the model.

### 2021-03-10 ST563-FP.Rmd
Added some model descriptions

@@
 ## Variable Selection
+
+## Basic Tree  
+The basic classification or regression tree is based on partitioning the data into subgroups using simple binary splitting.
+Initially, all objects are considered a single group.
+Then, the group is binarily split into two subgroups based on the criteria of a certain variable.
+In the classification setting, we classify the observations in a specific region with majority vote, while in the context of regression the mean of the group is used for prediction.
+We want to grow the tree as big as we can, and then prune it back using cost-complexity pruning.
+This is done to not overfit the data, but pruning increases the bias.
+The pruning parameter needs to be tuned, which is done automatically in the `caret` package.
+
+The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods like the random forest.
+Additionally, the basic classification or regression tree is not very computationally expensive, unlike the random forest, boosting, and the support vector machine.
+We used the `caret` and `rpart` packages in R to fit our trees.
+
+## Random Forest  
+The random forest model is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results.
+Many bootstrap samples are created with replacement and then a classification tree is fitted on each bootstrap sample with a random subset of the predictors.
+Once a prediction has been made by all of the bootstrapped trees, the final classification is based on majority vote of the bootstrap predictions.
+Similarly, the prediction in the regression context is an average of all of the predictions of the bootstrapped trees.
+
+The advantage for using an ensemble method over a regular classification or regression tree is that because there are many bootstrap samples being averaged together, there is less variance.
+This is similar to how the variance of the sample mean goes down as the sample size increases.
+Although it will probably increase our prediction accuracy, the random forest loses the interpretability that the basic trees have.
+Furthermore, the algorithm that is used to fit the random forest is very computationally expensive.
+
+To train our model, we used the `randomForest` and `caret` packages in R.
+The maximum number of predictors for a bootstrap sample as well as the number of trees in the forest are both parameters that need tuning.
+Again, we will use the "best" tune, automatically given to us by the `caret` package.
+
+## Support Vector Machine  
+A support vector machine is a type of classification rule where it essentially maximizes the margin between groups by choosing the "line" with the widest "margin", but allowing for error.
+We put "line" in quotation marks because with higher-dimensional data, the "line" is actually a hyper-plane-thing.
+Similarly, the "margin" doesn't actually exist the in sense of a clean margin between two things, since we are allowing for some sort of error.
+The support vectors are the points closest to the middle that carry the most weight when classifying, since they are the most prone to error and are deciding factors of where the classification line could go.
+We used a radial kernel function to deal with the non-linearity and higher-dimensions.
+
+The support vector machine is pretty effective in higher-dimensional spaces, but doesn't perform very well with very large data sets with lots of noise.
+If we were to look at hotels with millions of records, a support vector machine may not be suitable.
+We fit our SVM using the `kernlab` and `caret` packages in R.
+
+## K-Nearest Neighbors (KNN)  
+K-NN is a "model free" approach, meaning that it doesn't assume any probability model on the data.
+Given an observation, $\boldsymbol{x}$, we want to find the k training observations that are "closest" to $\boldsymbol{x}$, and then either classify the new observation using majority vote among these k neighbors or predict the new observation by averaging the closest k neighbors.
+We define "closeness" based on Euclidean distance in our model.
+As this is the case, we need to center and scale the data, so that different scales of measurement are kept constant.
+
+The K-nearest neighbors model is relatively intuitive and simple and has no underlying assumptions, making it a good model to consider.
+In our classification, we are only using it for a binary classification, but it is also very easy to extend the K-NN model to multiple classes.
+On the other hand, the K-NN algorithm is very sensitive to outliers, high dimensional data, and imbalanced data.
+
+The number (k) of closest neighbors is a parameter that needs tuning. We used the `class` and `caret` packages in R to fit this model. Based on our cross-validation, we used a k value of 5.
+
+## Logistic Regression  
+The logistic regression model uses the "logit" function $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$.
+Using it for binary classification, we round the fitted values either up or down to 1 or 0.
+The logistic regression function is defined as
+$$
+\ln \left( \frac{p(x)}{1-p(x)} \right) = x'\beta, \qquad \text{where }p(x) = (1+e^{-x'\beta})
+$$
+
+The logistic regression model assumes that each observation is independent, that there is little or no multicollinearity among the predictors, and that there is a linear relationship between the predictor variables and log odds.
+This differs from the typical regression model, where the residuals must be normally distributed and have constant variance.
+
+To fit the logistic regression model, we used the `glm()` function in base R.
+The advantages to using a logistic regression model are similar to those for the basic classification tree: it is not computationally expensive and easy to interpret the results.
+On the other hand, one of the major drawbacks to using a logistic regression model is that it is subject to the distributional assumptions on the data and errors mentioned above.
+If our assumptions are broken, our predictions may not be very reliable.
+
+## LASSO Regression
+
+## Boosting
+
+## Multiple Linear Regression

### 2021-03-11 ST563-FP.Rmd
added description for LASSO regression

@@
 
+## Cross Validation
 As mentioned in the introduction, our focus is on the classification accuracy of machine learning methods.
@@
 ## LASSO Regression
+LASSO, which stands for Least Absolute Shrinkage and Selection Operator works to shrink beta coefficient estimators towards zero due to a penalty.
+This is different from traditional shrinkage methods since the form of the LASSO penalty actually forces some of the coefficients to exactly zero.
+This helps achieve variable selection and yields sparser models, similar to subset selection.
+In particular, the LASSO estimate minimizes:
+$$
+\sum_{i=1}^n\left( y_i - \beta_0 - \sum_{j=1}^p\beta_jx_{ij}\right)^2 + \lambda \sum_{j=1}^p|\beta_j| = RSS + \lambda \sum_{j=1}^p|\beta_j|
+$$
+where $\lambda$ is the penalty parameter, with p different predictors.
+
+The main benefits of using LASSO are the variable selection property, the computational efficiency, and the interpretability.
+Plus, as with any regularization method, it can help to avoid overfitting with a very flexible model.
+On the other hand, some of the pitfalls of LASSO are that the selected variables could be highly biased, correlation among predictors is not very well dealt with, and the feature selection may not be stable over different bootstrapped samples.
+
+To fit the LASSO regression model, we used the `glmnet` package in R.
 

### 2021-03-11 ST563-FP.Rmd
started boosted tree description

@@
 
-## Boosting
+## Boosted Tree
+Boosting is the process of slowly training trees to avoid overfitting.
+The general approach is as follows: the trees are grown sequentially, where each subsequent tree is grown on a modified version of the original data, while the predictions are continuously updated as the trees are grown.
+The model has three parameters that need tuning:
 
+  * $B$ - the number of trees
+  * $\lambda$ - the shrinkage parameter, which is a small positive number that controls the rate at which the boosting learns
+  * $d$ - the number of splits in each tree, which controls the complexity of the boosted ensemble
+
+The main benefit of using boosting is to reduce overfitting. It often yields the best results in terms of the ensemble learning methods. 
 ## Multiple Linear Regression

### 2021-03-11 ST563-FP.Rmd
finished boosting description

@@
 
-The main benefit of using boosting is to reduce overfitting. It often yields the best results in terms of the ensemble learning methods. 
+The main benefit of using boosting is to reduce overfitting.
+It often yields the best results in terms of the ensemble learning methods.
+Like the random forest model, since boosting is an ensemble learning technique, it will have a lower variance than other methods.
+Also, there is no need to scale down variables or create any dummy variables for categorical variables.
+All that said, boosting is not without its shortcomings; it does not do well with high dimensionality, it's not very interpretable (since it is a sort of blackbox algorithm), and it is relatively expensive computationally.
+
 ## Multiple Linear Regression

### 2021-03-11 ST563-FP.Rmd
added description for MLR

@@
 # Methods  
-## Data Processing & Partitioning 
+## Data Processing
 ```{r partition}
@@
 ## Multiple Linear Regression
+Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
+The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
+Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
+The formula for a multiple linear regression model with $p$ predictor variables is as follows:
+$$
+\begin{aligned}
+&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
+& \quad \text{where for } i = n \text{ observations}, \\
+& \quad y_i = \text{dependent variable (response)} \\
+& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
+& \quad \beta_0 = \text{y-intercept} \\
+& \quad \beta_j = \text{slope coefficients for each variable} \\
+& \quad \epsilon_i = \text{residuals (error term of model)} \\
+\end{aligned}
+$$
+
+The multiple linear regression model is subject to the following assumptions about the data:
+  * Linear relationship between response and predictors
+  * No collinearity between predictor variables
+  * $y_i$'s are iid
+  * $\epsilon_i \sim N(0, \sigma^2)$
+
+The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
+The coefficients can be computed efficiently and they are have clear and easy interpretations.
+It is also possible to make inference on any of the variables or linear combination of variables.
+The most signficiant drawback of using the multiple linear regression model is that, like the logistic regression model, it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.

### 2021-03-11 ST563-FP.Rmd
added in some code for some of the models. Still need to figure out what variables go where...

@@
 For each wine, various features of the wine were measured, including a measurement of quality, given by an expert wine taster.
-Our objective it to use the different features to predict the quality of the wine.
+Our objective is to use the different features to predict the quality of the wine using different machine learning techniques.
 
@@
 
-Fortunately, the `wine` data set that we worked with was very tidy; there was no missing values.
+Fortunately, the `wine` dataset that we worked with was very tidy; there was no missing values.
 We didn't standardize the data initially, but when training the support vector machine as well as the k-nearest neighbors models we centered and scaled the data.
@@
 Since we used the `caret` package in R to fit all of our models, we used the tunes of the parameters that were deemed "best" by the `train()` function.
+
+```{r CV}
+# Specify the cross-validation method (5-fold CV)
+trctrl <- trainControl(method = "cv", number = 5)
+```
 
@@
 We used the `caret` and `rpart` packages in R to fit our trees.
+```{r tree, echo=FALSE}
+#### REGRESSION ####
+# Fit basic regression tree model
+tree.reg <- train(form, data = trn, method = "rpart",
+                  trControl = trctrl)
+
+# Obtain RMSE and results
+tree.reg.pred <- predict(tree.reg, newdata = tst)
+tree.reg.rmse <- mean((tree.reg.pred - tst$quality)^2)
+
+# Print out tree visualization
+tree.reg.out <- rpart(form, data = trn, cp = 0.01798144)
+rpart.plot(tree.reg.out, main="Regression Tree")
+
+#### HOTEL 2 ####
+# Fit classification tree model
+tree.class <- train(form, data = trn, method = "rpart",
+                    trControl = trctrl)
+
+# Obtain confusion matrix and results for H2
+tree.class.pred <- predict(tree.class, newdata = tst)
+tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
+tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])
+
+# Print out tree visualization
+tree.class.out <- rpart(form, data = trn, cp = 0.01798144)
+rpart.plot(tree.class.out, main="Classification Tree")
+
+# Print out confusion matrix
+# list("Confusion Matrix for H1" = h1.tree.CM$table,
+#      "Confusion Matrix for H2" = h2.tree.CM$table)
+```
 
@@
 
+```{r rf, echo=FALSE}
+#### REGRESSION ####
+# Formula for random forest
+form.rf <-
+
+# Fit the regression random forest model 
+rf.reg <- train(form.rf, data = trn, method = "rf",
+                trControl = trctrl, importance = TRUE)
+
+# Obtain RMSE and results
+rf.reg.pred <- predict(rf.reg, newdata = tst)
+rf.reg.rmse <- mean((rf.reg.pred - tst$quality)^2)
+
+#### CLASSIFICATION ####
+# Fit the classification random forest model
+rf.class <- train(form.rf, data = trn, method = "rf",
+                trControl = trctrl, importance = TRUE)
+
+# Obtain confusion matrix and results
+rf.class.pred <- predict(rf.class, newdata = tst)
+rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
+rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])
+
+# Print out confusion matrix
+# list("Confusion Matrix for H1" = h1.rf.CM$table,
+#      "Confusion Matrix for H2" = h2.rf.CM$table)
+```
+
 ## Support Vector Machine  
@@
 We fit our SVM using the `kernlab` and `caret` packages in R.
+
+```{r svm}
+#### CLASSIFICATION ####
+# Define formula for SVM model
+form.svm <-
+
+# Fit support vector machine for H1
+svm.class <- train(form.svm, data = trn, method = "svmRadial", 
+                   preProcess = c("center", "scale"),
+                   trControl = trctrl, tuneLength = 10)
+
+# Obtain confusion matrix and results for H1
+svm.class.pred <- predict(svm.class, newdata = tst)
+svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
+svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])
+```
 
@@
 
+```{r knn}
+#### REGRESSION ####
+# Define formula for KNN model
+form.knn <-
+
+# Fit knn model for H1
+knn.reg <- train(form.knn, data = trn, method = "knn",
+                 trControl = trctrl, preProcess = c("center", "scale"))
+
+# Obtain RMSE and results
+knn.reg.pred <- predict(knn.reg, newdata = tst)
+knn.reg.rmse <- mean((knn.reg.pred - tst$quality)^2)
+
+#### CLASSIFICATION ####
+# Fit knn model for H2
+knn.class <- train(form.kn, data = trn, method = "knn",
+                   trControl=trctrl, preProcess = c("center", "scale"))
+
+# Obtain confusion matrix and results for H2
+knn.class.pred <- predict(knn.class, newdata = tst)
+knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
+knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])
+```
+
 ## Logistic Regression  
-The logistic regression model uses the "logit" function $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$.
+The logistic regression model uses the "logit" function, $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$.
 Using it for binary classification, we round the fitted values either up or down to 1 or 0.
@@
 The most signficiant drawback of using the multiple linear regression model is that, like the logistic regression model, it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
+
+# Results
+
+# Conclusion

### 2021-03-11 ST563-FP.Rmd
added code for the first few models added variable selection code

--- 2021-03-11 ST563-FP.Rmd
@@
 library(class) # fitting kNN model
+library(glmnet) # fitting LASSO
 
@@
 
+```{r CV}
+# Specify the cross-validation method (5-fold CV)
+trctrl <- trainControl(method = "cv", number = 5)
+```
+
 ## Variable Selection
+After running an initial random forest model on the entire dataset with selected predictors, we looked at the variable importance of the predictors to decide which variables we would use in our final model.
+Variable importance, in the context on machine learning, refers to how much a given model "uses" that variable to make accurate predictions.
+In other words, the more a model relies on a variable to make predictions, the more important it is for the model.
+```{r varImp, echo = FALSE}
+# Write preliminary model  
+model.class.pre <- class ~ .
+
+# Fit preliminary random forest on entire data set
+rf.class.pre <- randomForest(model.class.pre, data=wine, ntree=500, mtry=3, importance=TRUE)
+
+# Create data frame of variable importance
+df.class.pre <- rf.class.pre$importance %>% as.data.frame()
+X <- rownames(df.class.pre)
+
+# Sort the data frame and prep for plot
+VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
+VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
+VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)
+
+# Plot variable importance
+vimp.class <- ggplot(VIrf.pre.class.sort[-12,],aes(x=X,y=MeanDecreaseAccuracy))+
+              geom_bar(stat = 'identity', position = 'dodge', fill="lightblue")+
+              theme(axis.text.x = element_text(angle = 60, hjust = 1))+
+              ylab("Mean Decrease in Accuracy") + xlab("Predictor")+
+              ggtitle("Variable Importance Plot")
+vimp.class
+```
+
+```{r}
+# Get X matrix
+X <- model.matrix(quality ~ ., data = wine)[, -1]
+
+# Fit LASSO model and plot result
+lm.lasso <- glmnet::cv.glmnet(X, wine$quality, alpha = 1)
+lambda <- lm.lasso$lambda.min
+plot(lm.lasso)
+
+# Fit LASSO model with best tune
+lasso.best <- glmnet(X, wine$quality, alpha = 1)
+predict(lasso.best, s = lambda, type = "coefficients")
+
+# Create model and get summary output
+lm.sub <- regsubsets(quality ~ ., data = wine, nvmax = 10)
+smre <- lm.sub %>% summary()
+
+# Plot the Cp, BIC, and R^2 results in one plot
+par(mfrow = c(2,2))
+plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
+points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
+plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
+points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
+plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
+points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
+
+# Best subset selection
+best.betas <- coef(lm.sub, id = 6) %>% round(2)
+best.betas
+```
+
+```{r form}
+form.reg <- quality ~ volatile.acidity + chlorides + 
+            total.sulfur.dioxide + pH + sulphates + alcohol
+form.class <- class ~ alcohol + sulphates + volatile.acidity + density + total.sulfur.dioxide
+```
 
@@
 We used the `caret` and `rpart` packages in R to fit our trees.
+
+```{r tree, echo=FALSE}
+#### REGRESSION ####
+# Fit basic regression tree model
+tree.reg <- train(form.reg, data = trn, method = "rpart",
+                  trControl = trctrl)
+
+# Obtain RMSE and results
+tree.reg.pred <- predict(tree.reg, newdata = tst)
+tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))
+
+#### HOTEL 2 ####
+# Fit classification tree model
+tree.class <- train(form.class, data = trn, method = "rpart",
+                    trControl = trctrl)
+
+# Obtain confusion matrix and results for H2
+tree.class.pred <- predict(tree.class, newdata = tst)
+tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
+tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])
+
+# Print out regression tree visualization
+tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
+rpart.plot(tree.reg.out, main="Regression Tree")
+
+# Print out classification tree visualization
+tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
+rpart.plot(tree.class.out, main="Classification Tree")
+
+# Print out confusion matrix
+list("RMSE" = tree.reg.rmse,
+     "Class Results" = tree.class.res)
+```
 
@@
 
+```{r rf, echo=FALSE}
+#### REGRESSION ####
+# Fit the regression random forest model 
+rf.reg <- train(form.reg, data = trn, method = "rf",
+                trControl = trctrl, importance = TRUE)
+
+# Obtain RMSE and results
+rf.reg.pred <- predict(rf.reg, newdata = tst)
+rf.reg.rmse <- mean((rf.reg.pred - tst$quality)^2)
+
+#### CLASSIFICATION ####
+# Fit the classification random forest model
+rf.class <- train(form.class, data = trn, method = "rf",
+                trControl = trctrl, importance = TRUE)
+
+# Obtain confusion matrix and results
+rf.class.pred <- predict(rf.class, newdata = tst)
+rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
+rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])
+
+# Print out results in a list
+list("RMSE" = rf.reg.rmse,
+     "Class Results" = rf.class.res)
+```
+
 ## Support Vector Machine  
@@
 We fit our SVM using the `kernlab` and `caret` packages in R.
+
+```{r svm}
+#### CLASSIFICATION ####
+# Fit support vector machine for H1
+svm.class <- train(form.class, data = trn, method = "svmRadial", 
+                   preProcess = c("center", "scale"),
+                   trControl = trctrl, tuneLength = 10)
+
+# Obtain confusion matrix and results for H1
+svm.class.pred <- predict(svm.class, newdata = tst)
+svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
+svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])
+
+# Print out results in a list
+list("Results" = svm.class.res,
+     "Confusion Matrix" = svm.class.CM$table)
+```
 
@@
 
+```{r knn}
+#### REGRESSION ####
+# Fit knn model for H1
+knn.reg <- train(form.reg, data = trn, method = "knn",
+                 trControl = trctrl, preProcess = c("center", "scale"))
+
+# Obtain RMSE and results
+knn.reg.pred <- predict(knn.reg, newdata = tst)
+knn.reg.rmse <- mean((knn.reg.pred - tst$quality)^2)
+
+#### CLASSIFICATION ####
+# Fit knn model for H2
+knn.class <- train(form.class, data = trn, method = "knn",
+                   trControl=trctrl, preProcess = c("center", "scale"))
+
+# Obtain confusion matrix and results for H2
+knn.class.pred <- predict(knn.class, newdata = tst)
+knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
+knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])
+
+# Print out results in a list
+list("RMSE" = knn.reg.rmse,
+     "Class Results" = knn.class.res)
+```
+
 ## Logistic Regression  
-The logistic regression model uses the "logit" function $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$.
+The logistic regression model uses the "logit" function, $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$.
 Using it for binary classification, we round the fitted values either up or down to 1 or 0.
@@
 If our assumptions are broken, our predictions may not be very reliable.
+
+```{r logit}
+#### CLASSIFICATION ####
+# Fit logistic regression model
+glm.class <- glm(form.class, data = trn,  family = binomial(link = "logit"))
+
+# Obtain confusion matrix and results for H2
+glm.class.pred <- predict(glm.class, newdata = tst, type = "response")
+glm.class.pred <- glm.class.pred %>% round() %>% as.factor()
+glm.class.CM <- confusionMatrix(glm.class.pred, ifelse(tst$class == "low", 1, 0) %>% as.factor())
+glm.class.res <- c(glm.class.CM$overall[1], glm.class.CM$byClass[1:2])
+
+# Print out results in a list
+list("Results" = glm.class.res,
+     "Confusion Matrix" = glm.class.CM$table)
+```
 

### 2021-03-11 ST563-FP.Rmd
added code for boosting added code for LASSO

@@
 library(glmnet) # fitting LASSO 
+library(gbm) # fitting the boosted tree model
 
@@
   * `class`: fitting the k-nearest neighbor model
+  * `glmnet`: fitting the LASSO model
+  * `gbm`: fitting the boosted tree model
 
@@
 ```{r}
-# Get X matrix
-X <- model.matrix(quality ~ ., data = wine)[, -1]
-
-# Fit LASSO model and plot result
-lm.lasso <- glmnet::cv.glmnet(X, wine$quality, alpha = 1)
-lambda <- lm.lasso$lambda.min
-plot(lm.lasso)
-
-# Fit LASSO model with best tune
-lasso.best <- glmnet(X, wine$quality, alpha = 1)
-predict(lasso.best, s = lambda, type = "coefficients")
-
 # Create model and get summary output
@@
 best.betas <- coef(lm.sub, id = 6) %>% round(2)
-best.betas
 ```
@@
 # Print out regression tree visualization
-tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
-rpart.plot(tree.reg.out, main="Regression Tree")
+#tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
+#rpart.plot(tree.reg.out, main="Regression Tree")
 
 # Print out classification tree visualization
-tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
-rpart.plot(tree.class.out, main="Classification Tree")
+#tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
+#rpart.plot(tree.class.out, main="Classification Tree")
 
 # Print out confusion matrix
-list("RMSE" = tree.reg.rmse,
-     "Class Results" = tree.class.res)
+#list("RMSE" = tree.reg.rmse,
+#     "Class Results" = tree.class.res)
 ```
@@
 rf.reg.pred <- predict(rf.reg, newdata = tst)
-rf.reg.rmse <- mean((rf.reg.pred - tst$quality)^2)
+rf.reg.rmse <- sqrt(mean((rf.reg.pred - tst$quality)^2))
 
@@
 # Print out results in a list
-list("RMSE" = rf.reg.rmse,
-     "Class Results" = rf.class.res)
+#list("RMSE" = rf.reg.rmse,
+#     "Class Results" = rf.class.res)
 ```
@@
 # Print out results in a list
-list("Results" = svm.class.res,
-     "Confusion Matrix" = svm.class.CM$table)
+#list("Results" = svm.class.res,
+#     "Confusion Matrix" = svm.class.CM$table)
 ```
@@
 knn.reg.pred <- predict(knn.reg, newdata = tst)
-knn.reg.rmse <- mean((knn.reg.pred - tst$quality)^2)
+knn.reg.rmse <- sqrt(mean((knn.reg.pred - tst$quality)^2))
 
@@
 # Print out results in a list
-list("RMSE" = knn.reg.rmse,
-     "Class Results" = knn.class.res)
+#list("RMSE" = knn.reg.rmse,
+#     "Class Results" = knn.class.res)
 ```
@@
 # Print out results in a list
-list("Results" = glm.class.res,
-     "Confusion Matrix" = glm.class.CM$table)
+#list("Results" = glm.class.res,
+#     "Confusion Matrix" = glm.class.CM$table)
 ```
@@
 ## LASSO Regression
-LASSO, which stands for Least Absolute Shrinkage and Selection Operator works to shrink beta coefficient estimators towards zero due to a penalty.
+LASSO, which stands for Least Absolute Shrinkage and Selection Operator works to shrink beta coefficient estimators towards zero using a penalty term.
 This is different from traditional shrinkage methods since the form of the LASSO penalty actually forces some of the coefficients to exactly zero.
@@
 To fit the LASSO regression model, we used the `glmnet` package in R.
+```{r}
+#### REGRESSION ####
+# Fit LASSO to the data
+lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
+                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))
+
+lasso.reg.pred <- predict(lasso.reg, newdata = tst)
+lasso.reg.rmse <- sqrt(mean((lasso.reg.pred - tst$quality)^2))
+#lasso.reg.rmse
+```
 
@@
 All that said, boosting is not without its shortcomings; it does not do well with high dimensionality, it's not very interpretable (since it is a sort of blackbox algorithm), and it is relatively expensive computationally.
+
+```{r}
+#### REGRESSION ####
+# Fit the regression boosted tree model 
+boost.reg <- train(form.reg, data = trn, method = "gbm",
+                trControl = trctrl, verbose = FALSE)
+
+# Obtain RMSE and results
+boost.reg.pred <- predict(boost.reg, newdata = tst)
+boost.reg.rmse <- sqrt(mean((boost.reg.pred - tst$quality)^2))
+#boost.reg.rmse
+```
 

### 2021-03-11 ST563-FP.Rmd
added code for MLR model
fixed issue with kable outputs

@@
 library(gbm) # fitting the boosted tree model
+library(kableExtra) # making bolded headers for tables
 
@@
                   colClasses = c(rep("double", 7), rep("double", 4), "integer"))
+
+# Set seed for reproducibility
+set.seed(702)
 ```
@@
         "KNN Classification", "Logistic Regression")
-reg <- c("Regression Tree", "Boosted Tree", "Multiple Linear Regression (MLR)", 
+reg <- c("Boosted Tree", "Random Forest", "Multiple Linear Regression (MLR)", 
           "KNN Regression", "LASSO Regression")
-kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods")
+kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods") %>% 
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% 
+  kable_styling(latex_options = "hold_position")
 ```
@@
 ## Variable Selection
-After running an initial random forest model on the entire dataset with selected predictors, we looked at the variable importance of the predictors to decide which variables we would use in our final model.
-Variable importance, in the context on machine learning, refers to how much a given model "uses" that variable to make accurate predictions.
+Since we have two different tasks that we are trying to complete (classification and regression), we went about selecting the variables for each task differently.
+For the classification task, we ran an initial random forest model on the entire dataset with selected predictors.
+We then looked at the variable importance of the predictors to decide which variables we would use in our final models.
+Variable importance, within the context of machine learning, refers to how much a given model "uses" that variable to make accurate predictions.
 In other words, the more a model relies on a variable to make predictions, the more important it is for the model.
-```{r varImp, echo = FALSE}
+
+```{r class.select, fig.align="center", out.width='4in'}
 # Write preliminary model  
@@
 # Fit preliminary random forest on entire data set
-rf.class.pre <- randomForest(model.class.pre, data=wine, ntree=500, mtry=3, importance=TRUE)
+rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
+                             ntree = 500, mtry = 3, importance = TRUE)
 
@@
 # Plot variable importance
-vimp.class <- ggplot(VIrf.pre.class.sort[-12,],aes(x=X,y=MeanDecreaseAccuracy))+
-              geom_bar(stat = 'identity', position = 'dodge', fill="lightblue")+
-              theme(axis.text.x = element_text(angle = 60, hjust = 1))+
-              ylab("Mean Decrease in Accuracy") + xlab("Predictor")+
-              ggtitle("Variable Importance Plot")
+vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
+                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
+                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
+                ggtitle("Variable Importance") + 
+                geom_hline(yintercept = .005, lty = 2, colour = "darkblue")
 vimp.class
@@
 
-```{r}
+We decided to choose the variables with a variable importance over .005, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`.
+The fact that density proved to be an important variable for classification was relatively surprising.
+One would think that the wine quality would be determined based on its attributes that are detectable by the human senses.
+For this same reason, alcohol, acidity, and sulphates made sense in terms of variable importance.
+Sulphates are essential for the preservation of the wine, an overly- or underly-acidic wine may be considered poor, and the alcohol content has a connection to the fermentation process, where the must (aka the grape juice before it undergoes fermentation) actually becomes wine.
+
+When looking at the problem in terms of a regression context, we used best subsets selection for feature selection.
+Using the `leaps` library and `regsubsets` in R, we are able to compare all of the possible models from the given set of predictors.
+Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
+We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
+
+```{r reg.select}
 # Create model and get summary output
-lm.sub <- regsubsets(quality ~ ., data = wine, nvmax = 10)
+lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
 smre <- lm.sub %>% summary()
 
-# Plot the Cp, BIC, and R^2 results in one plot
+# Plot the Cp, BIC, R^2, RSS results in one plot
 par(mfrow = c(2,2))
@@
 points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
+plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
+points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
 
@@
 
+Although each of the three model criteria suggested a different number of predictor variables of a different number, they were all within the same region (between 6 and 8 predictors).
+Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
+This was the model favored by BIC, which gives the greatest penalty for complexity in models.
+The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
+All of the variables from the classification selection appear here with the exception of density.
+In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses.
+The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.
+
 ```{r form}
-form.reg <- quality ~ volatile.acidity + chlorides + 
-            total.sulfur.dioxide + pH + sulphates + alcohol
-form.class <- class ~ alcohol + sulphates + volatile.acidity + density + total.sulfur.dioxide
+# Write formula for regression models
+form.reg <- quality ~ volatile.acidity + chlorides + pH + 
+            total.sulfur.dioxide + sulphates + alcohol
+
+# Write formula for classification models
+form.class <- class ~ alcohol + sulphates + density +
+              volatile.acidity + total.sulfur.dioxide
 ```
@@
 
-```{r tree, echo=FALSE}
+A visualization of the classification tree can be found below.
+
+```{r tree, fig.align="center", out.height='4in', out.width='4in'}
 #### REGRESSION ####
 # Fit basic regression tree model
-tree.reg <- train(form.reg, data = trn, method = "rpart",
-                  trControl = trctrl)
+#tree.reg <- train(form.reg, data = trn, method = "rpart",
+#                  trControl = trctrl)
 
 # Obtain RMSE and results
-tree.reg.pred <- predict(tree.reg, newdata = tst)
-tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))
+#tree.reg.pred <- predict(tree.reg, newdata = tst)
+#tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))
 
-#### HOTEL 2 ####
+#### CLASSIFICATION ####
 # Fit classification tree model
@@
 # Print out classification tree visualization
-#tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
-#rpart.plot(tree.class.out, main="Classification Tree")
-
-# Print out confusion matrix
-#list("RMSE" = tree.reg.rmse,
-#     "Class Results" = tree.class.res)
+tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
+rpart.plot(tree.class.out, main="Classification Tree")
 ```
@@
 
-```{r rf, echo=FALSE}
+```{r rf}
 #### REGRESSION ####
@@
 To fit the LASSO regression model, we used the `glmnet` package in R.
-```{r}
+```{r lasso}
 #### REGRESSION ####
@@
 
-```{r}
+```{r boost}
 #### REGRESSION ####
@@
 
+```{r mlr}
+#### REGRESSION ####
+lm.reg <- train(form.reg, data = trn, method = "lm",
+                trControl = trctrl)
+
+# Obtain RMSE and results for linear model
+lm.reg.pred <- predict(lm.reg, newdata = tst)
+lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
+#lm.reg.rmse
+```
+
 # Results
 
+```{r create.df}
+# Create data frame output for regression models
+df.reg <- rbind.data.frame(boost.reg.rmse, rf.reg.rmse,
+                           lm.reg.rmse, knn.reg.rmse, 
+                           lasso.reg.rmse) %>% round(4)
+df.reg$method <- reg
+colnames(df.reg) <- c(LETTERS[1:2])
+df.reg <- df.reg %>% select("B", "A")
+colnames(df.reg) <- c("Regression Method", "Root MSE")
+
+# Create data frame output for classification models
+df.class <- rbind.data.frame(tree.class.res, rf.class.res, 
+                             svm.class.res, knn.class.res, 
+                             glm.class.res) %>% round(4)
+df.class$method <- cl
+colnames(df.class) <- c("A", "B", "C", "D")
+df.class <- df.class %>% select("D", "A", "B", "C")
+colnames(df.class) <- c("Classification Method", "Accuracy", "Sensitivity", "Specificity")
+```
+
+```{r reg.out}
+# Smash tables next to each other using kable
+kable(df.reg, caption = "Regression Model Results", format = "latex") %>% 
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% kable_styling(latex_options = "HOLD_position")
+```
+
+```{r class.out}
+kable(df.class, caption = "Classification Model Results", format = "latex") %>% 
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
+  kable_styling(latex_options = "HOLD_position")
+```
+
 # Conclusion

### 2021-03-12 ST563-FP.Rmd
added appendix started results section

@@
 ```{r methods}
+# Make table of methods
 cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
@@
 #### REGRESSION ####
-# Fit knn model for H1
+# Fit knn regression model
 knn.reg <- train(form.reg, data = trn, method = "knn",
@@
 #### CLASSIFICATION ####
-# Fit knn model for H2
+# Fit knn classification model
 knn.class <- train(form.class, data = trn, method = "knn",
@@
 
-# Obtain confusion matrix and results for H2
+# Obtain confusion matrix and results
 knn.class.pred <- predict(knn.class, newdata = tst)
@@
 
+Based on the output, the best model for both classification and regression was the random forest.
+This was not a surprise, since the ensemble methods tend to outperform regular methods.
+This is further exemplified by the lower-than-average root MSE of the boosted tree model.
+The other models all did about the same in terms of prediction accuracy.
+For classification, all of the non-aforementioned methods were in the low 70s in terms of prediction accuracy, while the non-ensemble regression models had root MSEs right around 0.65.
+
 # Conclusion
+
+# Appendix
+```{r, eval=FALSE, echo=TRUE}
+#### SETUP STUFF ####
+# Set default chunk options
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+
+# Load necessary libraries
+library(tidyverse) # so many things
+library(knitr) # making nice tables
+library(caret) # training models & CV
+library(leaps) # subset selection
+library(corrplot) # visualize correlation
+library(rmarkdown) # for output documents
+library(rpart) # fitting classification tree
+library(rpart.plot) # visualization of classification tree
+library(randomForest) # fitting random forest model
+library(kernlab) # fitting SVM
+library(class) # fitting kNN model
+library(glmnet) # fitting LASSO 
+library(gbm) # fitting the boosted tree model
+library(kableExtra) # making bolded headers for tables
+
+# Read in data set
+wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
+                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))
+
+# Set seed for reproducibility
+set.seed(702)
+
+#### METHODS TABLE ####
+# Make table of methods
+cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
+        "KNN Classification", "Logistic Regression")
+reg <- c("Boosted Tree", "Random Forest", "Multiple Linear Regression (MLR)", 
+          "KNN Regression", "LASSO Regression")
+kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods") %>% 
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% 
+  kable_styling(latex_options = "hold_position")
+
+#### DATA PARTITIONING ####
+# Prepare data for classification
+wine$class <- ifelse(wine$quality <= 5, "low", "high") %>% as.factor()
+
+# Create Data Partition
+ind <- createDataPartition(wine$quality, p = .75, list = FALSE)
+
+# Create train/test data for H1
+trn <- wine[ind,]
+tst <- wine[-ind,]
+
+#### CROSS VALIDATIONN ####
+# Specify the cross-validation method (5-fold CV)
+trctrl <- trainControl(method = "cv", number = 5)
+
+#### VARIABLE SELECTION ####
+# Write preliminary model  
+model.class.pre <- class ~ .
+
+# Fit preliminary random forest on entire data set
+rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
+                             ntree = 500, mtry = 3, importance = TRUE)
+
+# Create data frame of variable importance
+df.class.pre <- rf.class.pre$importance %>% as.data.frame()
+X <- rownames(df.class.pre)
+
+# Sort the data frame and prep for plot
+VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
+VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
+VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)
+
+# Plot variable importance
+vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
+                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
+                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
+                ggtitle("Variable Importance") + 
+                geom_hline(yintercept = .005, lty = 2, colour = "darkblue")
+vimp.class
+
+# Create model and get summary output
+lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
+smre <- lm.sub %>% summary()
+
+# Plot the Cp, BIC, R^2, RSS results in one plot
+par(mfrow = c(2,2))
+plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
+points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
+plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
+points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
+plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
+points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
+plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
+points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
+
+# Best subset selection
+best.betas <- coef(lm.sub, id = 6) %>% round(2)
+
+#### WRITE FORMULAS ####
+# Write formula for regression models
+form.reg <- quality ~ volatile.acidity + chlorides + pH + 
+            total.sulfur.dioxide + sulphates + alcohol
+
+# Write formula for classification models
+form.class <- class ~ alcohol + sulphates + density +
+              volatile.acidity + total.sulfur.dioxide
+
+#### BASIC CLASSIFICATION TREE ####
+# Fit classification tree model
+tree.class <- train(form.class, data = trn, method = "rpart",
+                    trControl = trctrl)
+
+# Obtain confusion matrix and results for H2
+tree.class.pred <- predict(tree.class, newdata = tst)
+tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
+tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])
+
+# Print out classification tree visualization
+tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
+rpart.plot(tree.class.out, main="Classification Tree")
+
+#### RANDOM FOREST ####
+# Fit the regression random forest model 
+rf.reg <- train(form.reg, data = trn, method = "rf",
+                trControl = trctrl, importance = TRUE)
+
+# Obtain RMSE and results
+rf.reg.pred <- predict(rf.reg, newdata = tst)
+rf.reg.rmse <- sqrt(mean((rf.reg.pred - tst$quality)^2))
+
+# Fit the classification random forest model
+rf.class <- train(form.class, data = trn, method = "rf",
+                trControl = trctrl, importance = TRUE)
+
+# Obtain confusion matrix and results
+rf.class.pred <- predict(rf.class, newdata = tst)
+rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
+rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])
+
+#### SUPPORT VECTOR MACHINE ####
+# Fit support vector machine for H1
+svm.class <- train(form.class, data = trn, method = "svmRadial", 
+                   preProcess = c("center", "scale"),
+                   trControl = trctrl, tuneLength = 10)
+
+# Obtain confusion matrix and results for H1
+svm.class.pred <- predict(svm.class, newdata = tst)
+svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
+svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])
+
+#### KNN ####
+# Fit knn regression model
+knn.reg <- train(form.reg, data = trn, method = "knn",
+                 trControl = trctrl, preProcess = c("center", "scale"))
+
+# Obtain RMSE and results
+knn.reg.pred <- predict(knn.reg, newdata = tst)
+knn.reg.rmse <- sqrt(mean((knn.reg.pred - tst$quality)^2))
+
+# Fit knn classification model
+knn.class <- train(form.class, data = trn, method = "knn",
+                   trControl=trctrl, preProcess = c("center", "scale"))
+
+# Obtain confusion matrix and results
+knn.class.pred <- predict(knn.class, newdata = tst)
+knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
+knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])
+
+#### LOGISTIC REGRESSION ####
+# Fit logistic regression model
+glm.class <- glm(form.class, data = trn, family = binomial(link = "logit"))
+
+# Obtain confusion matrix and results for H2
+glm.class.pred <- predict(glm.class, newdata = tst, type = "response")
+glm.class.pred <- glm.class.pred %>% round() %>% as.factor()
+glm.class.CM <- confusionMatrix(glm.class.pred, ifelse(tst$class == "low", 1, 0) %>% as.factor())
+glm.class.res <- c(glm.class.CM$overall[1], glm.class.CM$byClass[1:2])
+
+#### LASSO ####
+# Fit LASSO to the data
+lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
+                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))
+
+lasso.reg.pred <- predict(lasso.reg, newdata = tst)
+lasso.reg.rmse <- sqrt(mean((lasso.reg.pred - tst$quality)^2))
+
+#### BOOSTING ####
+# Fit the regression boosted tree model 
+boost.reg <- train(form.reg, data = trn, method = "gbm",
+                trControl = trctrl, verbose = FALSE)
+
+# Obtain RMSE and results
+boost.reg.pred <- predict(boost.reg, newdata = tst)
+boost.reg.rmse <- sqrt(mean((boost.reg.pred - tst$quality)^2))
+
+#### MLR ####
+lm.reg <- train(form.reg, data = trn, method = "lm",
+                trControl = trctrl)
+
+# Obtain RMSE and results for linear model
+lm.reg.pred <- predict(lm.reg, newdata = tst)
+lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
+
+#### OUTPUT RESULTS ####
+# Create data frame output for regression models
+df.reg <- rbind.data.frame(boost.reg.rmse, rf.reg.rmse,
+                           lm.reg.rmse, knn.reg.rmse, 
+                           lasso.reg.rmse) %>% round(4)
+df.reg$method <- reg
+colnames(df.reg) <- c(LETTERS[1:2])
+df.reg <- df.reg %>% select("B", "A")
+colnames(df.reg) <- c("Regression Method", "Root MSE")
+
+# Create data frame output for classification models
+df.class <- rbind.data.frame(tree.class.res, rf.class.res, 
+                             svm.class.res, knn.class.res, 
+                             glm.class.res) %>% round(4)
+df.class$method <- cl
+colnames(df.class) <- c("A", "B", "C", "D")
+df.class <- df.class %>% select("D", "A", "B", "C")
+colnames(df.class) <- c("Classification Method", "Accuracy", "Sensitivity", "Specificity")
+
+# Smash tables next to each other using kable
+kable(df.reg, caption = "Regression Model Results", format = "latex") %>% 
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
+  kable_styling(latex_options = "HOLD_position")
+kable(df.class, caption = "Classification Model Results", format = "latex") %>% 
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
+  kable_styling(latex_options = "HOLD_position")
+```

### 2021-03-12 ST563-FP.Rmd
finsihed results section finished future research section finsihed limitations section

@@
 
-# Methods  
+## Data Exploration
+The first thing that I did with the data upon reading it in was take the numeric variables and create pairwise scatterplots.
+This way, we can get a feel for what variables are related as well as what the correlation structure looks like, since variables that are very related may cause problems moving forward.
+
+```{r pairs}
+# Create pairwise scatterplots
+pairs(wine[,c(2,5,7,8,9,10,11)])
+```
+Based on the pairwise scatterplots, we suspect that pH may be linearly correlated with density, sulphates, and alcohol.
+This leads me to believe that there is some relationship between the chemical properties of the wines.
+The following correlation plot gives a numeric summary of the linear relationship between the variables.
+
+```{r corrplot, echo = FALSE, fig.align='center', out.height='4in', out.width='4in'}
+# Create visualization of correlation structure
+p.mat <- cor.mtest(wine[,c(2,5,7,8,9,10,11)])$p
+corrplot(cor(wine[,c(2,5,7,8,9,10,11)]),
+         method = "color", type = "upper", order = "hclust", 
+         number.cex = .7, addCoef.col = "black",
+         tl.col = "black", tl.srt = 90,
+         p.mat = p.mat, sig.level = 0.01, insig = "blank", 
+         diag = T)
+```
+
+The only linear correlations that are on the higher side are between alcohol and density, which makes sense because the density of alcohol is slightly less than water.
+So as the concentration of alcohol gets higher, the density will naturally decrease.
+Similarly, the sulphates and chlorides seem to have a slightly positive linear relationship, while pH and density have a slightly negative linear relationship.
+
 ## Data Processing
@@
 
-Fortunately, the `wine` dataset that we worked with was very tidy; there was no missing values.
-We didn't standardize the data initially, but when training the support vector machine as well as the k-nearest neighbors models we centered and scaled the data.
+Fortunately, the `wine` dataset that we worked with was very tidy; there were no missing values.
+We didn't standardize the data initially, but when training the support vector machine and the k-nearest neighbors models we centered and scaled the data.
 
+# Methods
 ## Cross Validation
@@
 To fit the LASSO regression model, we used the `glmnet` package in R.
+
 ```{r lasso}
@@
 The multiple linear regression model is subject to the following assumptions about the data:
+
   * Linear relationship between response and predictors
@@
 
+The fact that the random forest so drastically outperformed the basic classification tree did not come as a surprise, since as an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
+We suspect that other ensemble methods, such as bagging or boosting, may also be able to get higher prediction accuracies.
+One of the main disadvantages to using ensemble techniques like the random forest is the fact that they tend to be very computationally intensive.
+The basic classification tree, logistic regression, and KNN models took significantly less time to run than the random forest and the SVM.
+If we wanted to run an ensemble method on a very large data set with a lot of features, it would be a computational nightmare.
+
+\newpage
+
 # Conclusion
 
+## Limitations  
+The multiple linear regression and logistic regression models were both subject to certain distributional and model assumptions, which were not explicitly tested or checked.
+Since the objective of this project was to predict rather than make inference, we did explicitly check the model assumptions.
+
+All of the wines in the dataset come from wineries in Portugal, and although they are from different wineries, we worry that the results may not be generalizable to wine production globally.
+Similarly, will these prediction results hold when it comes to white wine, rose, or sparkling wine?
+Given this variability over location and type of wine, what can we say about the winemaking industry as a whole?
+
+Some of the wine regions have more strict standards for the designations of wine quality.
+For example, regions like Bordeaux and Champagne in France have some of the highest standards in the world for their wines, which are much different than somewhere like Greece or New Zealand.
+How accurate will are predictions be when there are different standards and how might we account for that variation?
+
+## Further Research  
+Since this study solely looks at the prediction accuracy of machine learning techniques for red wines, the natural extension of the analysis would to verify that the results hold constant for white wines as well.
+The University of California, Irvine Machine Learning database has data on white wines, so that would be a good place to start.
+
+The characteristics of any given wine are often derived from different geological and/or climatological
+conditions.
+For example, a wine that comes from a region close enough to the water to get a sea breeze may have a hint of saltiness in its aroma or taste.
+Similarly, a grape coming from a relatively young grove might give the wine different features than if the grove were hundreds of years old.
+It would be interesting to see how different weather conditions, terroirs, plant ages, and slopes affect the outcome of the wine in terms of quality.
+
+\newpage
+
 # Appendix  
-```{r, eval=FALSE, echo=TRUE}
+```{r appendix, eval=FALSE, echo=TRUE}
 #### SETUP STUFF ####

### 2021-03-12 ST563-FP.Rmd
save

@@
 
-```{r pairs}
+```{r pairs, fig.align="center", out.width='4in', out.height='3.5in'}
 # Create pairwise scatterplots
@@
 # Conclusion  
+Hotels often suffer losses in revenue due to room vacancies that result from cancellations.
+Hotels can implement stringent cancellation policies as well as overbooking procedures to try and mitigate this loss in revenue.
+As analytics solutions are becoming more and more popular in the agriculture industry, we sought to apply machine learning learning models in an attempt to find .
+Hotel managers could use the results of a machine learning algorithm to take action based on knowledge of forecasted cancellations or demand.
+
+Based on the results of our study, we were able to predict the cancellation of a given room with about 75-80% accuracy.
+Although this is a *good* result considering the non-information rate of 54%, it is by no means excellent.
+These results indicate that machine learning algorithms could be a good technique for predicting cancellations, but that there is room for improvement.
+With a more formal, industry expert-guided variable selection process, we believe that higher prediction accuracies can be achieved.
+Machine learning models could potentially be combined with other techniques to help optimize revenue from bookings.
 

### 2021-03-12 ST563-FP.Rmd
3:00AM on Friday, March 12, 2021 I am officially done with the final project

--- 2021-03-12 ST563-FP.Rmd
@@
     toc_depth: 1
+fontsize: 12pt
 ---
@@
 There has been a push towards the implementation of different analytics solutions within the agriculture industry, especially in the realms of crop health and yield.
-These sort of analyses can help farmers to reduce waste and improve profits.
-In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is multifaceted.
+These sort of analyses can help farmers reduce waste and improve profits.
+In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is lengthy and multifaceted.
 Winemakers can work to optimize revenue by analyzing their input and output (crops and wines).
@@
 
-\newpage
-
 # Data
@@
 The data set is composed of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine.
-For each wine, various features of the wine were measured, including a measurement of quality, given by an expert wine taster.
+For each wine, various physical and chemical features of the wine were measured, including a measurement of quality given by an expert wine taster.
 Our objective is to use the different features to predict the quality of the wine using different machine learning techniques.
@@
 ## Data Exploration
-The first thing that I did with the data upon reading it in was take the numeric variables and create pairwise scatterplots.
+The first thing that we did with the data upon reading it in was take the numeric variables and create pairwise scatterplots.
 This way, we can get a feel for what variables are related as well as what the correlation structure looks like, since variables that are very related may cause problems moving forward.
 
-```{r pairs}
+```{r pairs, fig.align="center", out.width='4in', out.height='3.5in'}
 # Create pairwise scatterplots
@@
 Based on the pairwise scatterplots, we suspect that pH may be linearly correlated with density, sulphates, and alcohol.
-This leads me to believe that there is some relationship between the chemical properties of the wines.
+This leads us to believe that there is some relationship between the chemical properties of the wines.
 The following correlation plot gives a numeric summary of the linear relationship between the variables.
@@
 We didn't standardize the data initially, but when training the support vector machine and the k-nearest neighbors models we centered and scaled the data.
+We also created the variable for classification, transforming the quality variable into a binary "low" or "high" response as mentioned in the introduction.
 
@@
 ## Cross Validation
-As mentioned in the introduction, our focus is on the classification accuracy of machine learning methods.
+Again, our focus is on the classification accuracy of machine learning methods.
 In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the models.
-We split the data into a training and test data set in order to later evaluate the model's prediction accuracy.
+We split the data into a training and test data set in order to later evaluate the model's prediction accuracy or root mean squared error (RMSE).
 For this project we used a 75/25 split, training the data on the 75% and testing the trained models on the withheld 25%.
@@
 We then looked at the variable importance of the predictors to decide which variables we would use in our final models.
-Variable importance, within the context of machine learning, refers to how much a given model "uses" that variable to make accurate predictions.
+Within the context of machine learning, variable importance refers to how much a given model "uses" that variable to make accurate predictions.
 In other words, the more a model relies on a variable to make predictions, the more important it is for the model.
@@
                 ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
-                ggtitle("Variable Importance") + 
-                geom_hline(yintercept = .005, lty = 2, colour = "darkblue")
+                ggtitle("Variable Importance")
 vimp.class
@@
 
-We decided to choose the variables with a variable importance over .005, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`.
+We decided to choose the five variables with the most variable importance, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`.
 The fact that density proved to be an important variable for classification was relatively surprising.
@@
 When looking at the problem in terms of a regression context, we used best subsets selection for feature selection.
-Using the `leaps` library and `regsubsets` in R, we are able to compare all of the possible models from the given set of predictors.
+Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors.
 Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
@@
 
-Although each of the three model criteria suggested a different number of predictor variables of a different number, they were all within the same region (between 6 and 8 predictors).
+Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
 Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
@@
 Initially, all objects are considered a single group.
-Then, the group is binarily split into two subgroups based on the criteria of a certain variable.
+Then, the group is repeatedly binarily split into two subgroups based on the criteria of a certain variable.
 In the classification setting, we classify the observations in a specific region with majority vote, while in the context of regression the mean of the group is used for prediction.
@@
 
-  * $B$ - the number of trees
-  * $\lambda$ - the shrinkage parameter, which is a small positive number that controls the rate at which the boosting learns
-  * $d$ - the number of splits in each tree, which controls the complexity of the boosted ensemble
+  * $B$: the number of trees
+  * $\lambda$: the shrinkage parameter, which is a small positive number that controls the rate at which the boosting learns
+  * $d$: the number of splits in each tree, which controls the complexity of the boosted ensemble
 
@@
 kable(df.reg, caption = "Regression Model Results", format = "latex") %>% 
-  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% kable_styling(latex_options = "HOLD_position")
+  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
+  kable_styling(latex_options = "HOLD_position")
 ```
@@
 
-The fact that the random forest so drastically outperformed the basic classification tree did not come as a surprise, since as an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
-We suspect that other ensemble methods, such as bagging or boosting, may also be able to get higher prediction accuracies.
+The fact that the random forest and boosted tree models so drastically outperformed the basic classification tree did not come as a surprise, since as an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
+We suspect that other ensemble methods, such as bagging, may also be able to get higher prediction accuracies.
 One of the main disadvantages to using ensemble techniques like the random forest is the fact that they tend to be very computationally intensive.
-The basic classification tree, logistic regression, and KNN models took significantly less time to run than the random forest and the SVM.
+The basic classification tree, logistic regression, and KNN models took significantly less time to run than the random forest and the SVM, but they were less accurate in their predictions.
 If we wanted to run an ensemble method on a very large data set with a lot of features, it would be a computational nightmare.
 
-\newpage
+# Conclusion  
+With the recent push for quantitative solutions in industries like agriculture, there has been an emphasis on using machine learning to solve a variety of problems requiring prediction and inference.
+In particular, the viticulture industry stands to gain a lot from such methods, since historically it has not interacted much with the world of analytics.
+In this study, we went through a number of different machine learning techniques, assessing prediction accuracy in both classification and regression contexts.
 
-# Conclusion
+Based on the results of our study, we were able to predict the cancellation of a given room with about 70-80% accuracy.
+Although this is a *good* result considering the non-information rate of 53%, it is by no means excellent.
+These results indicate that machine learning algorithms could be a good technique when trying to predict wine quality, but that there is room for improvement.
+With a more formal, industry expert-guided variable selection process, we believe that higher prediction accuracies can be achieved.
+Machine learning models could potentially be combined with other techniques to add value to winemakers.
 
@@
 
-# Appendix  
-```{r appendix, eval=FALSE, echo=TRUE}
+```{r appendix, eval=FALSE, echo=FALSE}
 #### SETUP STUFF ####

### 2021-04-13 ST563-FP.Rmd
Jackie Update KNN Classification

@@
 library(kableExtra) # making bolded headers for tables
+library(class) #for KNN
 
@@
 
-# Create Data Partition
-ind <- createDataPartition(wine$quality, p = .75, list = FALSE)
+# Create Test and Training Sets
+ind <- sample(length(wine$class), 0.8*length(wine$class))
 
@@
 # Fit knn classification model
-knn.class <- train(form.class, data = trn, method = "knn",
-                   trControl=trctrl, preProcess = c("center", "scale"))
+trn.class <- wine$class[ind]
+trn.knn <- trn[,-13]
+tst.knn <- tst[,-13]
+knn.class <- knn(trn.knn, tst.knn, trn.class, k=5)
 
-# Obtain confusion matrix and results
-knn.class.pred <- predict(knn.class, newdata = tst)
-knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
-knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])
+# Obtain confusion matrix and misclassification rate
+table(knn.class, tst$class)
+(37+39)/(37+39+140+104)
 

### 2021-04-15 ST563-FP.Rmd
Jackie SVM update

--- 2021-03-12 ST563-FP.Rmd
@@
 library(kableExtra) # making bolded headers for tables
+library(class) #for KNN
+library(e1071) #for support vector machine
 
@@
 ## Support Vector Machine  
-A support vector machine is a type of classification rule where it essentially maximizes the margin between groups by choosing the "line" with the widest "margin", but allowing for error.
-We put "line" in quotation marks because with higher-dimensional data, the "line" is actually a hyper-plane-thing.
-Similarly, the "margin" doesn't actually exist the in sense of a clean margin between two things, since we are allowing for some sort of error.
-The support vectors are the points closest to the middle that carry the most weight when classifying, since they are the most prone to error and are deciding factors of where the classification line could go.
-We used a radial kernel function to deal with the non-linearity and higher-dimensions.
-
-The support vector machine is pretty effective in higher-dimensional spaces, but doesn't perform very well with very large data sets with lots of noise.
-If we were to look at hotels with millions of records, a support vector machine may not be suitable.
-We fit our SVM using the `kernlab` and `caret` packages in R.
+A support vector machine (SVM) is a classification method, which maximizes the margin between groups by choosing the "line" with the widest "margin", while still allowing for error.
+Line in quotation marks because with higher-dimensional data, the "line" is a hyperplane.
+Similarly, the "margin" does not exist in the sense of a clean margin between two, since the model is allowing for some error.
+The support vectors are the points closest to the middle, these carry the most weight in classification.
+This is because they are most prone to error and are deciding factors of where the classification line could go.
+The SVM was built using cross validation, where multiple SVMs were built using different values for the cost parameter.
+The `tune()` function builds these models and the best is selected,  in this case the cost parameter of 0.01 was the best.
+For this data, the SVM correctly predicted all observations.
+If the data were much larger, one would expect the model to be less effective due to the extra noise.
 
@@
 # Fit support vector machine for H1
-svm.class <- train(form.class, data = trn, method = "svmRadial", 
-                   preProcess = c("center", "scale"),
-                   trControl = trctrl, tuneLength = 10)
-
+svm.class <- tune(svm, class~ . ,data=trn ,kernel ="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1,5,10,100) ))
+best.mod <- svm.class$best.model
 # Obtain confusion matrix and results for H1
-svm.class.pred <- predict(svm.class, newdata = tst)
-svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
-svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])
+svm.class.pred <- predict(best.mod, newdata = tst)
+table(svm.class.pred, tst$class)
 

### 2021-04-15 Tree_Rong.R
rong's code

@@ -1,0 +1,144 @@
+## ST563 Final project
+## Author: Rong Huang
+## 4/15/2021
+
+library(tidyverse)
+library(caret)
+require(tree)
+require(randomForest)
+require(gbm)
+
+# Read in data set
+wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
+                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))
+# Set seed for reproducibility
+set.seed(702)
+## Data Processing
+# Prepare data for classification
+wine$class <- ifelse(wine$quality <= 5, "low", "high") %>% as.factor()
+
+# Create Data Partition
+ind <- createDataPartition(wine$quality, p = .75, list = FALSE)
+# Create train/test data for H1
+train <- wine[ind,]
+test <- wine[-ind,]
+
+#--------------------
+#    Desion Tree
+#--------------------
+
+#----- classification -----
+# 1. full size decision tree
+tree.class=tree(class~.-quality,train)
+plot(tree.class)
+text(tree.class, pretty=0)
+tree.class.pred=predict(tree.class,test,type='class')
+tree.class.train_err=mean(train$class!=predict(tree.class,type='class'))
+tree.class.test_err=mean(test$class!=tree.class.pred)
+
+# 2. pruned decision tree 
+cv.tree.class=cv.tree(tree.class,FUN=prune.misclass)
+plot(cv.tree.class)
+prune.tree.class=prune.misclass(tree.class,best=8)
+plot(prune.tree.class);text(prune.tree.class,pretty=0)
+prune.tree.class.pred=predict(prune.tree.class,test,type='class')
+prune.tree.class.train_err=mean(train$class!=predict(prune.tree.class,type='class'))
+prune.tree.class.test_err=mean(test$class!=prune.tree.class.pred)
+
+#----- regression -----
+# 1. full size decision tree
+tree.reg=tree(quality~.-class,train)
+plot(tree.reg)
+text(tree.reg, pretty=0)
+tree.reg.pred=predict(tree.reg,test)
+tree.reg.train_rmse=sqrt(mean((train$quality-predict(tree.reg))^2))
+tree.reg.test_rmse=sqrt(mean((test$quality-tree.reg.pred)^2))
+
+# 2. pruned decision tree 
+cv.tree.reg=cv.tree(tree.reg)
+plot(cv.tree.reg)
+prune.tree.reg=prune.tree(tree.reg,best=8)
[…]

### 2021-04-19 DRAFT.Rmd
Add draft .Rmd file and output pdf Took pertinent sections from the ST563-FP.Rmd file and made some small changes to adapt code.

--- 2021-04-15 ST563-FP.Rmd
@@
 title: "ST563 Final Project"
-author: "Mana Azizsoltani"
+author: "Final Project Group 6"
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 # Make table of methods
-cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
-        "KNN Classification", "Logistic Regression")
-reg <- c("Boosted Tree", "Random Forest", "Multiple Linear Regression (MLR)", 
-          "KNN Regression", "LASSO Regression")
+cl <- c("Classification Tree", "Random Forest")
+reg <- c("Multiple Linear Regression (MLR)", "Random Forest")
 kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods") %>% 
@@
   * `randomForest`: fitting random forest models
-  * `kernlab`: fitting the support vector machine
-  * `class`: fitting the k-nearest neighbor model
-  * `glmnet`: fitting the LASSO model
-  * `gbm`: fitting the boosted tree model
 
@@
 We will then repeat this process over 5 folds of the data, averaging the results.
-
-The tree-based, LASSO, and K-nearest neighbors models require parameters to be tuned (more on those later).
-Since we used the `caret` package in R to fit all of our models, we used the tunes of the parameters that were deemed "best" by the `train()` function.
 
@@
 
-## Basic Tree  
-The basic classification or regression tree is based on partitioning the data into subgroups using simple binary splitting.
+## Multiple Linear Regression
+Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
+The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
+Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
+The formula for a multiple linear regression model with $p$ predictor variables is as follows:
+$$
+\begin{aligned}
+&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
+& \quad \text{where for } i = n \text{ observations}, \\
+& \quad y_i = \text{dependent variable (response)} \\
+& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
+& \quad \beta_0 = \text{y-intercept} \\
+& \quad \beta_j = \text{slope coefficients for each variable} \\
+& \quad \epsilon_i = \text{residuals (error term of model)} \\
+\end{aligned}
+$$
+
+The multiple linear regression model is subject to the following assumptions about the data:
+
+  * Linear relationship between response and predictors
+  * No collinearity between predictor variables
+  * $y_i$'s are iid
+  * $\epsilon_i \sim N(0, \sigma^2)$
+
+The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
+The coefficients can be computed efficiently and they are have clear and easy interpretations.
+It is also possible to make inference on any of the variables or linear combination of variables.
+The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
+
+```{r mlr}
+#### REGRESSION ####
+lm.reg <- train(form.reg, data = trn, method = "lm",
+                trControl = trctrl)
+
+# Obtain RMSE and results for linear model
+lm.reg.pred <- predict(lm.reg, newdata = tst)
+lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
+#lm.reg.rmse
+```
+
+## Classification Tree  
+The basic classification tree is based on partitioning the data into subgroups using simple binary splitting.
 Initially, all objects are considered a single group.
 Then, the group is repeatedly binarily split into two subgroups based on the criteria of a certain variable.
-In the classification setting, we classify the observations in a specific region with majority vote, while in the context of regression the mean of the group is used for prediction.
+In the classification setting, we classify the observations in a specific region with majority vote.
 We want to grow the tree as big as we can, and then prune it back using cost-complexity pruning.
@@
 
-The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods like the random forest.
-Additionally, the basic classification or regression tree is not very computationally expensive, unlike the random forest, boosting, and the support vector machine.
+The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
+Additionally, the basic classification tree is not very computationally expensive, unlike the random forest, boosting, and support vector machine.
 We used the `caret` and `rpart` packages in R to fit our trees.
@@
 
-## Support Vector Machine  
-A support vector machine (SVM) is a classification method, which maximizes the margin between groups by choosing the "line" with the widest "margin", while still allowing for error.
-Line in quotation marks because with higher-dimensional data, the "line" is a hyperplane.
-Similarly, the "margin" does not exist in the sense of a clean margin between two, since the model is allowing for some error.
-The support vectors are the points closest to the middle, these carry the most weight in classification.
-This is because they are most prone to error and are deciding factors of where the classification line could go.
-The SVM was built using cross validation, where multiple SVMs were built using different values for the cost parameter.
-The `tune()` function builds these models and the best is selected,  in this case the cost parameter of 0.01 was the best.
-For this data, the SVM correctly predicted all observations.
-If the data were much larger, one would expect the model to be less effective due to the extra noise.
-
-```{r svm}
-#### CLASSIFICATION ####
-# Fit support vector machine for H1
-svm.class <- tune(svm, class~ . ,data=trn ,kernel ="linear", ranges=list(cost=c(0.001, 0.01, 0.1, 1,5,10,100) ))
-best.mod <- svm.class$best.model
-# Obtain confusion matrix and results for H1
-svm.class.pred <- predict(best.mod, newdata = tst)
-table(svm.class.pred, tst$class)
-
-# Print out results in a list
-#list("Results" = svm.class.res,
-#     "Confusion Matrix" = svm.class.CM$table)
-```
-
-## K-Nearest Neighbors (KNN)  
-K-NN is a "model free" approach, meaning that it doesn't assume any probability model on the data.
-Given an observation, $\boldsymbol{x}$, we want to find the k training observations that are "closest" to $\boldsymbol{x}$, and then either classify the new observation using majority vote among these k neighbors or predict the new observation by averaging the closest k neighbors.
-We define "closeness" based on Euclidean distance in our model.
-As this is the case, we need to center and scale the data, so that different scales of measurement are kept constant.
-
-The K-nearest neighbors model is relatively intuitive and simple and has no underlying assumptions, making it a good model to consider.
-In our classification, we are only using it for a binary classification, but it is also very easy to extend the K-NN model to multiple classes.
-On the other hand, the K-NN algorithm is very sensitive to outliers, high dimensional data, and imbalanced data.
-
-The number (k) of closest neighbors is a parameter that needs tuning. We used the `class` and `caret` packages in R to fit this model. Based on our cross-validation, we used a k value of 5.
-
-```{r knn}
-#### REGRESSION ####
-# Fit knn regression model
-knn.reg <- train(form.reg, data = trn, method = "knn",
-                 trControl = trctrl, preProcess = c("center", "scale"))
-
-# Obtain RMSE and results
-knn.reg.pred <- predict(knn.reg, newdata = tst)
-knn.reg.rmse <- sqrt(mean((knn.reg.pred - tst$quality)^2))
-
-#### CLASSIFICATION ####
-# Fit knn classification model
-trn.class <- wine$class[ind]
-trn.knn <- trn[,-13]
-tst.knn <- tst[,-13]
-knn.class <- knn(trn.knn, tst.knn, trn.class, k=5)
-
-# Obtain confusion matrix and misclassification rate
-table(knn.class, tst$class)
-(37+39)/(37+39+140+104)
-
-# Print out results in a list
-#list("RMSE" = knn.reg.rmse,
-#     "Class Results" = knn.class.res)
-```
-
-## Logistic Regression  
-The logistic regression model uses the "logit" function, $log(\frac{p}{1-p})$, which links the mean to the linear form of the regression model, $\boldsymbol{X\beta}$.
-Using it for binary classification, we round the fitted values either up or down to 1 or 0.
-The logistic regression function is defined as
-$$
-\ln \left( \frac{p(x)}{1-p(x)} \right) = x'\beta, \qquad \text{where }p(x) = (1+e^{-x'\beta})
-$$
-
-The logistic regression model assumes that each observation is independent, that there is little or no multicollinearity among the predictors, and that there is a linear relationship between the predictor variables and log odds.
-This differs from the typical regression model, where the residuals must be normally distributed and have constant variance.
-
-To fit the logistic regression model, we used the `glm()` function in base R.
-The advantages to using a logistic regression model are similar to those for the basic classification tree: it is not computationally expensive and easy to interpret the results.
-On the other hand, one of the major drawbacks to using a logistic regression model is that it is subject to the distributional assumptions on the data and errors mentioned above.
-If our assumptions are broken, our predictions may not be very reliable.
-
-```{r logit}
-#### CLASSIFICATION ####
-# Fit logistic regression model
-glm.class <- glm(form.class, data = trn, family = binomial(link = "logit"))
-
-# Obtain confusion matrix and results for H2
-glm.class.pred <- predict(glm.class, newdata = tst, type = "response")
-glm.class.pred <- glm.class.pred %>% round() %>% as.factor()
-glm.class.CM <- confusionMatrix(glm.class.pred, ifelse(tst$class == "low", 1, 0) %>% as.factor())
-glm.class.res <- c(glm.class.CM$overall[1], glm.class.CM$byClass[1:2])
-
-# Print out results in a list
-#list("Results" = glm.class.res,
-#     "Confusion Matrix" = glm.class.CM$table)
-```
-
-## LASSO Regression
-LASSO, which stands for Least Absolute Shrinkage and Selection Operator works to shrink beta coefficient estimators towards zero using a penalty term.
-This is different from traditional shrinkage methods since the form of the LASSO penalty actually forces some of the coefficients to exactly zero.
-This helps achieve variable selection and yields sparser models, similar to subset selection.
-In particular, the LASSO estimate minimizes:
-$$
-\sum_{i=1}^n\left( y_i - \beta_0 - \sum_{j=1}^p\beta_jx_{ij}\right)^2 + \lambda \sum_{j=1}^p|\beta_j| = RSS + \lambda \sum_{j=1}^p|\beta_j|
-$$
-where $\lambda$ is the penalty parameter, with p different predictors.
-
-The main benefits of using LASSO are the variable selection property, the computational efficiency, and the interpretability.
-Plus, as with any regularization method, it can help to avoid overfitting with a very flexible model.
-On the other hand, some of the pitfalls of LASSO are that the selected variables could be highly biased, correlation among predictors is not very well dealt with, and the feature selection may not be stable over different bootstrapped samples.
-
-To fit the LASSO regression model, we used the `glmnet` package in R.
-
-```{r lasso}
-#### REGRESSION ####
-# Fit LASSO to the data
-lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
-                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))
-
-lasso.reg.pred <- predict(lasso.reg, newdata = tst)
-lasso.reg.rmse <- sqrt(mean((lasso.reg.pred - tst$quality)^2))
-#lasso.reg.rmse
-```
-
-## Boosted Tree
-Boosting is the process of slowly training trees to avoid overfitting.
-The general approach is as follows: the trees are grown sequentially, where each subsequent tree is grown on a modified version of the original data, while the predictions are continuously updated as the trees are grown.
-The model has three parameters that need tuning:
-
-  * $B$: the number of trees
-  * $\lambda$: the shrinkage parameter, which is a small positive number that controls the rate at which the boosting learns
-  * $d$: the number of splits in each tree, which controls the complexity of the boosted ensemble
-
-The main benefit of using boosting is to reduce overfitting.
-It often yields the best results in terms of the ensemble learning methods.
-Like the random forest model, since boosting is an ensemble learning technique, it will have a lower variance than other methods.
-Also, there is no need to scale down variables or create any dummy variables for categorical variables.
-All that said, boosting is not without its shortcomings; it does not do well with high dimensionality, it's not very interpretable (since it is a sort of blackbox algorithm), and it is relatively expensive computationally.
-
-```{r boost}
-#### REGRESSION ####
-# Fit the regression boosted tree model 
-boost.reg <- train(form.reg, data = trn, method = "gbm",
-                trControl = trctrl, verbose = FALSE)
-
-# Obtain RMSE and results
-boost.reg.pred <- predict(boost.reg, newdata = tst)
-boost.reg.rmse <- sqrt(mean((boost.reg.pred - tst$quality)^2))
-#boost.reg.rmse
-```
-
-## Multiple Linear Regression
-Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
-The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
-Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
-The formula for a multiple linear regression model with $p$ predictor variables is as follows:
-$$
-\begin{aligned}
-&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
-& \quad \text{where for } i = n \text{ observations}, \\
-& \quad y_i = \text{dependent variable (response)} \\
-& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
-& \quad \beta_0 = \text{y-intercept} \\
-& \quad \beta_j = \text{slope coefficients for each variable} \\
-& \quad \epsilon_i = \text{residuals (error term of model)} \\
-\end{aligned}
-$$
-
-The multiple linear regression model is subject to the following assumptions about the data:
-
-  * Linear relationship between response and predictors
-  * No collinearity between predictor variables
-  * $y_i$'s are iid
-  * $\epsilon_i \sim N(0, \sigma^2)$
-
-The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
-The coefficients can be computed efficiently and they are have clear and easy interpretations.
-It is also possible to make inference on any of the variables or linear combination of variables.
-The most signficiant drawback of using the multiple linear regression model is that, like the logistic regression model, it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
-
-```{r mlr}
-#### REGRESSION ####
-lm.reg <- train(form.reg, data = trn, method = "lm",
-                trControl = trctrl)
-
-# Obtain RMSE and results for linear model
-lm.reg.pred <- predict(lm.reg, newdata = tst)
-lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
-#lm.reg.rmse
-```
-
 # Results
@@
 # Create data frame output for regression models
-df.reg <- rbind.data.frame(boost.reg.rmse, rf.reg.rmse,
-                           lm.reg.rmse, knn.reg.rmse, 
-                           lasso.reg.rmse) %>% round(4)
+df.reg <- rbind.data.frame(lm.reg.rmse, rf.reg.rmse) %>% round(4)
 df.reg$method <- reg
@@
 # Create data frame output for classification models
-df.class <- rbind.data.frame(tree.class.res, rf.class.res, 
-                             svm.class.res, knn.class.res, 
-                             glm.class.res) %>% round(4)
+df.class <- rbind.data.frame(tree.class.res, rf.class.res) %>% round(4)
 df.class$method <- cl
@@
 This was not a surprise, since the ensemble methods tend to outperform regular methods.
-This is further exemplified by the lower-than-average root MSE of the boosted tree model.
-The other models all did about the same in terms of prediction accuracy.
-For classification, all of the non-aforementioned methods were in the low 70s in terms of prediction accuracy, while the non-ensemble regression models had root MSEs right around 0.65.
+As an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
+We suspect that other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracies.
 
-The fact that the random forest and boosted tree models so drastically outperformed the basic classification tree did not come as a surprise, since as an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
-We suspect that other ensemble methods, such as bagging, may also be able to get higher prediction accuracies.
 One of the main disadvantages to using ensemble techniques like the random forest is the fact that they tend to be very computationally intensive.
-The basic classification tree, logistic regression, and KNN models took significantly less time to run than the random forest and the SVM, but they were less accurate in their predictions.
+The basic classification tree and multiple linear regression model took significantly less time to run than the random forest and the SVM, but they were less accurate in their predictions.
 If we wanted to run an ensemble method on a very large data set with a lot of features, it would be a computational nightmare.
@@
 In particular, the viticulture industry stands to gain a lot from such methods, since historically it has not interacted much with the world of analytics.
-In this study, we went through a number of different machine learning techniques, assessing prediction accuracy in both classification and regression contexts.
+In this study, we went through various machine learning techniques, assessing prediction accuracy in both classification and regression contexts.
 
-Based on the results of our study, we were able to predict the cancellation of a given room with about 70-80% accuracy.
+Based on the results of our study, we were able to predict the high or low quality of a given wine with about 70-80% accuracy.
 Although this is a *good* result considering the non-information rate of 53%, it is by no means excellent.
@@
 ## Limitations  
-The multiple linear regression and logistic regression models were both subject to certain distributional and model assumptions, which were not explicitly tested or checked.
-Since the objective of this project was to predict rather than make inference, we did explicitly check the model assumptions.
+The multiple linear regression model is subject to certain distributional and model assumptions, which were not explicitly tested or checked.
+Since the objective of this project was to predict rather than make inference, we didn't explicitly check the model assumptions.
 
@@
 It would be interesting to see how different weather conditions, terroirs, plant ages, and slopes affect the outcome of the wine in terms of quality.
-
-\newpage
-
-```{r appendix, eval=FALSE, echo=FALSE}
-#### SETUP STUFF ####
-# Set default chunk options
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
-
-# Load necessary libraries
-library(tidyverse) # so many things
-library(knitr) # making nice tables
-library(caret) # training models & CV
-library(leaps) # subset selection
-library(corrplot) # visualize correlation
-library(rmarkdown) # for output documents
-library(rpart) # fitting classification tree
-library(rpart.plot) # visualization of classification tree
-library(randomForest) # fitting random forest model
-library(kernlab) # fitting SVM
-library(class) # fitting kNN model
-library(glmnet) # fitting LASSO 
-library(gbm) # fitting the boosted tree model
-library(kableExtra) # making bolded headers for tables
-
-# Read in data set
-wine <- read.csv2(file = "winequality-red.csv", header = TRUE, dec = ".",
-                  colClasses = c(rep("double", 7), rep("double", 4), "integer"))
-
-# Set seed for reproducibility
-set.seed(702)
-
-#### METHODS TABLE ####
-# Make table of methods
-cl <- c("Classification Tree", "Random Forest", "Support Vector Machine (SVM)",
-        "KNN Classification", "Logistic Regression")
-reg <- c("Boosted Tree", "Random Forest", "Multiple Linear Regression (MLR)", 
-          "KNN Regression", "LASSO Regression")
-kable(data.frame("Classification" = cl, "Regression" = reg), caption = "ML Methods") %>% 
-  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>% 
-  kable_styling(latex_options = "hold_position")
-
-#### DATA PARTITIONING ####
-# Prepare data for classification
-wine$class <- ifelse(wine$quality <= 5, "low", "high") %>% as.factor()
-
-# Create Data Partition
-ind <- createDataPartition(wine$quality, p = .75, list = FALSE)
-
-# Create train/test data for H1
-trn <- wine[ind,]
-tst <- wine[-ind,]
-
-#### CROSS VALIDATIONN ####
-# Specify the cross-validation method (5-fold CV)
-trctrl <- trainControl(method = "cv", number = 5)
-
-#### VARIABLE SELECTION ####
-# Write preliminary model  
-model.class.pre <- class ~ .
-
-# Fit preliminary random forest on entire data set
-rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
-                             ntree = 500, mtry = 3, importance = TRUE)
-
-# Create data frame of variable importance
-df.class.pre <- rf.class.pre$importance %>% as.data.frame()
-X <- rownames(df.class.pre)
-
-# Sort the data frame and prep for plot
-VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
-VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
-VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)
-
-# Plot variable importance
-vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
-                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
-                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
-                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
-                ggtitle("Variable Importance") + 
-                geom_hline(yintercept = .005, lty = 2, colour = "darkblue")
-vimp.class
-
-# Create model and get summary output
-lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
-smre <- lm.sub %>% summary()
-
-# Plot the Cp, BIC, R^2, RSS results in one plot
-par(mfrow = c(2,2))
-plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
-points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
-plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
-points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
-plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
-points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
-plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
-points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
-
-# Best subset selection
-best.betas <- coef(lm.sub, id = 6) %>% round(2)
-
-#### WRITE FORMULAS ####
-# Write formula for regression models
-form.reg <- quality ~ volatile.acidity + chlorides + pH + 
-            total.sulfur.dioxide + sulphates + alcohol
-
-# Write formula for classification models
-form.class <- class ~ alcohol + sulphates + density +
-              volatile.acidity + total.sulfur.dioxide
-
-#### BASIC CLASSIFICATION TREE ####
-# Fit classification tree model
-tree.class <- train(form.class, data = trn, method = "rpart",
-                    trControl = trctrl)
-
-# Obtain confusion matrix and results for H2
-tree.class.pred <- predict(tree.class, newdata = tst)
-tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
-tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])
-
-# Print out classification tree visualization
-tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
-rpart.plot(tree.class.out, main="Classification Tree")
-
-#### RANDOM FOREST ####
-# Fit the regression random forest model 
-rf.reg <- train(form.reg, data = trn, method = "rf",
-                trControl = trctrl, importance = TRUE)
-
-# Obtain RMSE and results
-rf.reg.pred <- predict(rf.reg, newdata = tst)
-rf.reg.rmse <- sqrt(mean((rf.reg.pred - tst$quality)^2))
-
-# Fit the classification random forest model
-rf.class <- train(form.class, data = trn, method = "rf",
-                trControl = trctrl, importance = TRUE)
-
-# Obtain confusion matrix and results
-rf.class.pred <- predict(rf.class, newdata = tst)
-rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
-rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])
-
-#### SUPPORT VECTOR MACHINE ####
-# Fit support vector machine for H1
-svm.class <- train(form.class, data = trn, method = "svmRadial", 
-                   preProcess = c("center", "scale"),
-                   trControl = trctrl, tuneLength = 10)
-
-# Obtain confusion matrix and results for H1
-svm.class.pred <- predict(svm.class, newdata = tst)
-svm.class.CM <- confusionMatrix(svm.class.pred, tst$class)
-svm.class.res <- c(svm.class.CM$overall[1], svm.class.CM$byClass[1:2])
-
-#### KNN ####
-# Fit knn regression model
-knn.reg <- train(form.reg, data = trn, method = "knn",
-                 trControl = trctrl, preProcess = c("center", "scale"))
-
-# Obtain RMSE and results
-knn.reg.pred <- predict(knn.reg, newdata = tst)
-knn.reg.rmse <- sqrt(mean((knn.reg.pred - tst$quality)^2))
-
-# Fit knn classification model
-knn.class <- train(form.class, data = trn, method = "knn",
-                   trControl=trctrl, preProcess = c("center", "scale"))
-
-# Obtain confusion matrix and results
-knn.class.pred <- predict(knn.class, newdata = tst)
-knn.class.CM <- confusionMatrix(knn.class.pred, tst$class)
-knn.class.res <- c(knn.class.CM$overall[1], knn.class.CM$byClass[1:2])
-
-#### LOGISTIC REGRESSION ####
-# Fit logistic regression model
-glm.class <- glm(form.class, data = trn, family = binomial(link = "logit"))
-
-# Obtain confusion matrix and results for H2
-glm.class.pred <- predict(glm.class, newdata = tst, type = "response")
-glm.class.pred <- glm.class.pred %>% round() %>% as.factor()
-glm.class.CM <- confusionMatrix(glm.class.pred, ifelse(tst$class == "low", 1, 0) %>% as.factor())
-glm.class.res <- c(glm.class.CM$overall[1], glm.class.CM$byClass[1:2])
-
-#### LASSO ####
-# Fit LASSO to the data
-lasso.reg <- train(form.reg, data = trn, method = "glmnet", trControl = trctrl, 
-                   tuneGrid = expand.grid(.alpha=1, .lambda=0.00175))
-
-lasso.reg.pred <- predict(lasso.reg, newdata = tst)
-lasso.reg.rmse <- sqrt(mean((lasso.reg.pred - tst$quality)^2))
-
-#### BOOSTING ####
-# Fit the regression boosted tree model 
-boost.reg <- train(form.reg, data = trn, method = "gbm",
-                trControl = trctrl, verbose = FALSE)
-
-# Obtain RMSE and results
-boost.reg.pred <- predict(boost.reg, newdata = tst)
-boost.reg.rmse <- sqrt(mean((boost.reg.pred - tst$quality)^2))
-
-#### MLR ####
-lm.reg <- train(form.reg, data = trn, method = "lm",
-                trControl = trctrl)
-
-# Obtain RMSE and results for linear model
-lm.reg.pred <- predict(lm.reg, newdata = tst)
-lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
-
-#### OUTPUT RESULTS ####
-# Create data frame output for regression models
-df.reg <- rbind.data.frame(boost.reg.rmse, rf.reg.rmse,
-                           lm.reg.rmse, knn.reg.rmse, 
-                           lasso.reg.rmse) %>% round(4)
-df.reg$method <- reg
-colnames(df.reg) <- c(LETTERS[1:2])
-df.reg <- df.reg %>% select("B", "A")
-colnames(df.reg) <- c("Regression Method", "Root MSE")
-
-# Create data frame output for classification models
-df.class <- rbind.data.frame(tree.class.res, rf.class.res, 
-                             svm.class.res, knn.class.res, 
-                             glm.class.res) %>% round(4)
-df.class$method <- cl
-colnames(df.class) <- c("A", "B", "C", "D")
-df.class <- df.class %>% select("D", "A", "B", "C")
-colnames(df.class) <- c("Classification Method", "Accuracy", "Sensitivity", "Specificity")
-
-# Smash tables next to each other using kable
-kable(df.reg, caption = "Regression Model Results", format = "latex") %>% 
-  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
-  kable_styling(latex_options = "HOLD_position")
-kable(df.class, caption = "Classification Model Results", format = "latex") %>% 
-  row_spec(0, bold=TRUE) %>% collapse_rows(columns = 1, latex_hline = "none") %>%
-  kable_styling(latex_options = "HOLD_position")
-```

### 2021-04-19 DRAFT.Rmd
added some stuff off the top of my head put the MLR model first instead of last couldn't find yun's code so I quickly put together some residual plots changed some of the wording and added some tuning plots

@@
 # Set default chunk options
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
+                      warning = FALSE, fig.align = "center")
 
@@
 library(randomForest) # fitting random forest model
-library(kernlab) # fitting SVM
-library(class) # fitting kNN model
-library(glmnet) # fitting LASSO 
-library(gbm) # fitting the boosted tree model
 library(kableExtra) # making bolded headers for tables
-library(class) #for KNN
-library(e1071) #for support vector machine
+library(gridExtra) # stitching graphs together
 
@@
 
-```{r pairs, fig.align="center", out.width='4in', out.height='3.5in'}
+```{r pairs, out.width='4in', out.height='3.5in'}
 # Create pairwise scatterplots
@@
 
-```{r corrplot, echo = FALSE, fig.align='center', out.height='4in', out.width='4in'}
+```{r corrplot, echo = FALSE, out.height='3.5in'}
 # Create visualization of correlation structure
@@
 
-```{r class.select, fig.align="center", out.width='4in'}
+```{r class.select, out.height='4in'}
 # Write preliminary model  
@@
 
-```{r reg.select}
+```{r reg.select, out.height='3.5in'}
 # Create model and get summary output
@@
 ## Multiple Linear Regression
+```{r mlr}
+#### REGRESSION ####
+lm.reg <- lm(form.reg, data = trn)
+
+# Obtain RMSE and results for linear model
+lm.reg.pred <- predict(lm.reg, newdata = tst)
+lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
+```
 Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
@@
 
+To assess these assumptions, we looked at various diagnostics plots:
+```{r lm.diag, out.height = '3.5in'}
+# Look at residual plot for nonlinearity
+regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
+                 aes(x = lm.reg$residuals, y = lm.reg$fitted)) + 
+            geom_point(position = 'jitter') + labs(x = "Residuals", y = "Fitted Values")
+
+# Look at standardized residual plot for outliers
+stdres <- ggplot(data = data.frame(x = 1:length(lm.reg$residuals), 
+                                   y = scale(lm.reg$residuals)), 
+                 aes(y = y, x = x)) + 
+            geom_point() + labs(x = "Index", y = "Standardized Residuals")
+
+# Look at a qqplot to assess normality
+ggqq <- ggplot(data.frame(lm.reg$residuals), aes(sample = lm.reg$residuals)) + 
+          stat_qq() + stat_qq_line(col = "blue", lty = 2) + 
+          labs(x = "Theoretical", y = "Sample")
+
+# Look at residuals vs leverage plot
+reslev <- ggplot(lm.reg, aes(x = .hat, y = .stdresid)) + 
+            geom_point(aes(size = .cooksd)) + 
+            stat_smooth(method = "lm") + 
+            labs( x = "Leverage", y = "Standardized Residuals") + 
+            scale_size_continuous("Cook's D", range = c(1,5)) +
+            theme_bw() + theme(legend.position = "right")
+
+# Stitch all the graphs together
+grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)
+```
+
+~ talk about residual plots here ~
+
 The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
@@
 The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
-
-```{r mlr}
-#### REGRESSION ####
-lm.reg <- train(form.reg, data = trn, method = "lm",
-                trControl = trctrl)
-
-# Obtain RMSE and results for linear model
-lm.reg.pred <- predict(lm.reg, newdata = tst)
-lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
-#lm.reg.rmse
-```
 
@@
 This is done to not overfit the data, but pruning increases the bias.
-The pruning parameter needs to be tuned, which is done automatically in the `caret` package.
+The pruning parameter needs to be tuned; below is a plot of the classification accuracy by pruning parameter, where .001 yields the highest accuracy.
+
+```{r tree, out.height='3in'}
+# Set seed for reprodicibility
+set.seed(702)
+
+# Fit classification tree model
+tree.class <- train(form.class, data = trn, method = "rpart",
+                    trControl = trctrl, tuneGrid = expand.grid(cp = seq(0, .02, .001)))
+
+# Plot the accuracy as a function of the tuning parameter
+plot(Accuracy ~ cp, data = tree.class$results, xlab = "Cp")
+points(which.max(tree.class$results$Accuracy),
+       tree.class$results$Accuracy[which.max(tree.class$results$Accuracy)],
+       pch = 8, col = "#CC0000")
+```
 
 The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
-Additionally, the basic classification tree is not very computationally expensive, unlike the random forest, boosting, and support vector machine.
+Additionally, the classification tree is not very computationally expensive, unlike the random forest model.
 We used the `caret` and `rpart` packages in R to fit our trees.
@@
 
-```{r tree, fig.align="center", out.height='4in', out.width='4in'}
-#### REGRESSION ####
-# Fit basic regression tree model
-#tree.reg <- train(form.reg, data = trn, method = "rpart",
-#                  trControl = trctrl)
-
-# Obtain RMSE and results
-#tree.reg.pred <- predict(tree.reg, newdata = tst)
-#tree.reg.rmse <- sqrt(mean((tree.reg.pred - tst$quality)^2))
-
-#### CLASSIFICATION ####
-# Fit classification tree model
-tree.class <- train(form.class, data = trn, method = "rpart",
-                    trControl = trctrl)
-
+```{r, out.heigh = '3.5in'}
 # Obtain confusion matrix and results for H2
@@
 
-# Print out regression tree visualization
-#tree.reg.out <- rpart(form.reg, data = trn, cp = 0.03061684)
-#rpart.plot(tree.reg.out, main="Regression Tree")
-
 # Print out classification tree visualization
-tree.class.out <- rpart(form.class, data = trn, cp = 0.01971326)
+tree.class.out <- rpart(form.class, data = trn, cp = 0.011)
 rpart.plot(tree.class.out, main="Classification Tree")
@@
 
-The advantage for using an ensemble method over a regular classification or regression tree is that because there are many bootstrap samples being averaged together, there is less variance.
-This is similar to how the variance of the sample mean goes down as the sample size increases.
-Although it will probably increase our prediction accuracy, the random forest loses the interpretability that the basic trees have.
-Furthermore, the algorithm that is used to fit the random forest is very computationally expensive.
+The following parameters need to be tuned in a random forest model:  
+  * `mtry`: the number of randomly-selected variables selected at a node split
+  * `ntree`: the number of trees to grow
 
-To train our model, we used the `randomForest` and `caret` packages in R.
-The maximum number of predictors for a bootstrap sample as well as the number of trees in the forest are both parameters that need tuning.
-Again, we will use the "best" tune, automatically given to us by the `caret` package.
+```{r rf, out.height='3.5in'}
+#### REGRESSION ####
+# Set seed for reproducibility
+set.seed(702)
 
-```{r rf}
-#### REGRESSION ####
 # Fit the regression random forest model 
 rf.reg <- train(form.reg, data = trn, method = "rf",
-                trControl = trctrl, importance = TRUE)
+                trControl = trctrl, importance = TRUE,
+                tuneGrid = expand.grid(mtry = 1:6))
 
@@
 rf.class <- train(form.class, data = trn, method = "rf",
-                trControl = trctrl, importance = TRUE)
+                trControl = trctrl, importance = TRUE,
+                tuneGrid = expand.grid(mtry = 1:5))
 
@@
 
-# Print out results in a list
-#list("RMSE" = rf.reg.rmse,
-#     "Class Results" = rf.class.res)
+# Plot the RMSE and accuracy as a function of mtry (ntree constant at 500)
+par(mfrow = c(1,2))
+plot(RMSE ~ mtry, data = rf.reg$results)
+points(which.min(rf.reg$results$RMSE), rf.reg$results$RMSE[which.min(rf.reg$results$RMSE)],
+       pch = 8, col = "#CC0000")
+plot(Accuracy ~ mtry, data = rf.class$results)
+points(which.max(rf.class$results$Accuracy),
+       rf.class$results$Accuracy[which.max(rf.class$results$Accuracy)],
+       pch = 8, col = "#CC0000")
 ```
+
+Since `ntree` will plateau with enough trees, we tuned the `mtry` in both the regression and classification settings.
+For regression, the optimal `mtry` value is `r rf.reg$bestTune[1,1]`, while for classification it is  `r rf.class$bestTune[1,1]`.
+Keep in mind that the random forest is equivalent to bagging when the number of randomly selected features is equal to the total number of variables in the model.
+In both of the above plots, this is represented by the maximum displayed value of `mtry`.
+
+The advantage for using an ensemble method over a regular classification or regression tree is that because there are many bootstrap samples being averaged together, there is less variance.
+This is similar to how the variance of the sample mean goes down as the sample size increases.
+Although it will probably increase our prediction accuracy, the random forest loses the interpretability that the basic trees have.
+Furthermore, the algorithm that is used to fit the random forest is very computationally expensive.
 
@@
 One of the main disadvantages to using ensemble techniques like the random forest is the fact that they tend to be very computationally intensive.
-The basic classification tree and multiple linear regression model took significantly less time to run than the random forest and the SVM, but they were less accurate in their predictions.
+The basic classification tree and multiple linear regression model took significantly less time to run than the random forest, but they were less accurate in their predictions.
 If we wanted to run an ensemble method on a very large data set with a lot of features, it would be a computational nightmare.

### 2021-04-20 DRAFT.Rmd
Jackie Updates

@@
 library(gridExtra) # stitching graphs together
+library(tree)
 
@@
 # Introduction
+
 There has been a push towards the implementation of different analytics solutions within the agriculture industry, especially in the realms of crop health and yield.
@@
 Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality.
+The question this study is most interested in is to determine if there is a way to predict the quality of a wine without having to use taste testers.
+The team is hypothesizing that both the classifation and regression methods can accurately describe the quality of wine.
 
-In this study, our objective is to try and predict the quality of wine using a variety of statistical learning techniques.
-We will look at this problem from a regression standpoint as well as a classification standpoint.
-Using regression, we will try and predict the exact rating (0 to 10) of a particular wine.
-On the other hand, when we look at classification, we will try to classify a particular wine as either good or bad, taking wines rated from 0-5 as low quality wines and wines rated from 6-10 as high quality wines.
-We will be assessing the prediction accuracy of the following methods:
+In this study, the objective is to model and predict the quality of wine using a variety of statistical learning techniques.
+This will be done by performing both regression and classification methods, with regression ranking the quality of the wines on a scale of zero to ten, with zero being the lowest quality and ten being the highest quality.
+The classification models will be classifying the wines into two categories, good and bad, with wines in the bad category having a score of zero to five, and wines in the good category having a quality score between six and ten.
+The assessment of the prediction accuracy will be performed by the following methods:
 
@@
 # Data
-The data set used is the `Wine Quality` dataset from the UCI Machine Learning Repository.
-The data set is composed of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine.
-For each wine, various physical and chemical features of the wine were measured, including a measurement of quality given by an expert wine taster.
-Our objective is to use the different features to predict the quality of the wine using different machine learning techniques.
+The data set used is the Wine Quality dataset from the UCI Machine Learning Repository.
+The data set is comprised of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine.
+For each wine, various physical and chemical features of each wine was measured, including a measurement of quality given by an expert wine taster.
+The objective is to use the different features to predict the quality of the wine using different machine learning techniques.
 
@@
 Each entry in the dataset represents the different metrics of the following attributes of a single type of wine.
-
   * **fixed acidity**: the quantity of fixed acids found in the wines.
-The predominant fixed acids found in wines are tartaric, malic, citric, and succinic, all of which come from the grapes with the exception of the succinic acid, which comes from the yeast in fermentation.
+The predominant fixed acids found in wines are tartaric, malic, citric, and succinic, all of which come from the grapes except for the succinic acid, which comes from the yeast in fermentation.
   * **volatile acidity**: the steam distillable acids present in wine, primarily acetic acid but also lactic, formic, butyric, and propionic acids.
@@
 ## Data Exploration
-The first thing that we did with the data upon reading it in was take the numeric variables and create pairwise scatterplots.
-This way, we can get a feel for what variables are related as well as what the correlation structure looks like, since variables that are very related may cause problems moving forward.
+The first thing that was done to the data upon reading it was to use the numeric variables and create pairwise scatterplots.
+This way, one can obtain a greater understanding of how the variables relate to one another, as well as if there is any correlation structure and what it looks like.
+If the variables are related, there may be problems moving forward, so it is important to know these relationships before the analysis is performed.
 
@@
 ```
-Based on the pairwise scatterplots, we suspect that pH may be linearly correlated with density, sulphates, and alcohol.
+Based on the pairwise scatterplots, one can conclude that pH may be linearly correlated with density, sulphates, and alcohol.
 This leads us to believe that there is some relationship between the chemical properties of the wines.
+here also appears to be a relationship between density and alcohol.
 The following correlation plot gives a numeric summary of the linear relationship between the variables.
@@
 
-The only linear correlations that are on the higher side are between alcohol and density, which makes sense because the density of alcohol is slightly less than water.
-So as the concentration of alcohol gets higher, the density will naturally decrease.
+The only linear correlations that are notably high are between alcohol and density.
+This makes sense because the density of alcohol is slightly less than water, meaning that as the concentration of alcohol gets higher, the density will naturally decrease.
 Similarly, the sulphates and chlorides seem to have a slightly positive linear relationship, while pH and density have a slightly negative linear relationship.
+
+### Good Wine Qualities v. Bad Wine Qualities
+
+After a split in the data was performed to pull out the wines classified as good and the wines classified as bad.
+Further analysis was performed on each group to see if any of the variables were significantly different between the two groups.
+
+```{r good v bad}
+good.wine <- subset(wine, quality > 5)
+bad.wine <- subset(wine, quality <= 5)
+summary(good.wine)
+summary(bad.wine)
+```
+The two groups were similar in most of the variables.
+Two notable differences are that the wines classified as "good" had a lower mean value than "bad" wines for total sulfur dioxide.
+Good wines had a mean of 39, which was significantly lower than the mean of the "bad" wines at 54.6.
+Another difference to note is the difference in the mean for alcohol.
+The good wines had a mean alcohol content at 10.8, while the bad wines had a lower mean alcohol content around 9.9.
+There were also differences in fixed acidity, with the bad wines having a lower mean.
+There were also several variables that don't seem to change between the two groups including pH, residual sugar, chlorides, and density.
 
@@
 
-Fortunately, the `wine` dataset that we worked with was very tidy; there were no missing values.
-We didn't standardize the data initially, but when training the support vector machine and the k-nearest neighbors models we centered and scaled the data.
-We also created the variable for classification, transforming the quality variable into a binary "low" or "high" response as mentioned in the introduction.
+Fortunately, the dataset was very tidy; there were no missing values.
+The data was first split into a training and test data set.
+The training data set uses 80% of the observations and the test set contains the remaining 20%.
+The models are built on the training set and evaluated on the test set.
+This is important so that the models are not overfit.
+A variable was then created for classification, transforming the quality variable into a binary "low" or "high" response as mentioned in the introduction.
 
-# Methods
+# Classification Models
+## Classification Tree
+
+THIS SECTION NEEDS TO BE EDITED
+
+The basic classification tree is based on partitioning the data into subgroups using simple binary splitting.
+Initially, all objects are considered a single group.
+Then, the group is repeatedly split into two subgroups based on the criteria of a certain variable.
+In the classification setting, an observation is classified in a specific region with majority vote.
+Trees are grown as big as possible and then pruned using cost-complexity pruning.
+This is done to ensure the data is not being overfit.
+The tradeoff here is that pruning increases the bias.
+The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
+Additionally, the classification tree is not very computationally expensive, unlike the random forest model.
+
+```{r tree, out.height='3in'}
+# Set seed for reprodicibility
+set.seed(702)
+
+# 1. full size decision tree
+tree.class<-tree(class~.-quality,trn)
+plot(tree.class); text(tree.class, pretty=0)
+tree.class.pred<-predict(tree.class,tst,type='class')
+tree.class.train_err<-mean(trn$class!=predict(tree.class,type='class'))
+tree.class.test_err<-mean(tst$class!=tree.class.pred)
+tree.class.train_err
+tree.class.test_err
+```
+The resulting classification tree creates splits on 4 variables, Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity.
+There are 10 terminal nodes.
+The tree has a test error rate of 0.297.
+This test rate is fairly low, but as mentioned before, if one conducts pruning on the tree, the error rate can decrease and a more accurate model is created.
+
+Pruning is performed on the tree using cross validation, the tree is built many times with a different number of splits and the tree that produces the smallest error rate is selected.
+Cross validation on this model shows that the best model is the model built with 8 splits.
+The resulting pruned model will now have 8 terminal nodes and still uses the four variables in the un-pruned tree.
+```{r pruned tree}
+# 2. pruned decision tree 
+cv.tree.class<-cv.tree(tree.class,FUN=prune.misclass)
+plot(cv.tree.class)
+prune.tree.class<-prune.misclass(tree.class,best=8)
+plot(prune.tree.class);text(prune.tree.class,pretty=0)
+prune.tree.class.pred<-predict(prune.tree.class,tst,type='class')
+prune.tree.class.train_err<-mean(trn$class!=predict(prune.tree.class,type='class'))
+prune.tree.class.test_err<-mean(tst$class!=prune.tree.class.pred)
+prune.tree.class.test_err
+```
+The pruned tree produces a test error rate to the original tree with a test MSE of 0.3.
+The pruned tree would be selected using the concept of Occam's Razor, where the simplest model is the best model choice.
+
+Using a Random Forest tree can often improve upon the error rate of a simple classification tree.
+While the test MSE was small in the classification tree, the team believed the error rate could be decreases in using a more complex tree based method.
+
+## Random Forest Classification
+The random forest model is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees.
+Many bootstrap samples are created with replacement and then a classification tree is fitted on each bootstrap sample with a random subset of the predictors.
+Once a prediction has been made by the bootstrapped trees, the final classification is based on majority vote of the bootstrap predictions.
+Similarly, the prediction in the regression context is an average of the predictions of the bootstrapped trees.
+
+The advantage for using an ensemble method over a regular classification or regression tree is that because there are many bootstrap samples being averaged together, there is less variance.
+This is similar to how the variance of the sample mean goes down as the sample size increases.
+Although it will probably increase our prediction accuracy, the random forest loses the interpretability that the basic trees have.
+Furthermore, the algorithm that is used to fit the random forest is very computationally expensive.
+
+The first thing that should be done in building a boosted tree is cross validation.
+Here the model is built many times with the number of predictors included from one to all predictors.
+The test MSE is evaluated for each and the smallest value is selected.
+A plot of these is included below, the tree that produced the smallest error rate was the tree built with two variables, so that was the tree selected through cross validation.
+```{r rf classification CV}
+#----- classification -----
+set.seed(702)
+test.err <- double(11)
+for (mtry in 1:11){
+  fit <- randomForest(class~.-quality, trn,mtry=mtry)
+  pred <- predict(fit,tst)
+  test.err[mtry] <- mean(tst$class != pred)
+}
+matplot(1:mtry,test.err,pch=19,col='red',type='b',
+        ylab='Mean Square Error')
+```
+
+```{r rf class models}
+rf.class <- randomForest(class~.-quality,trn,mtry=2)
+rf.class.pred <- predict(rf.class,tst)
+rf.class.train_err <- mean(trn$class!=predict(rf.class))
+rf.class.test_err <- mean(tst$class!=rf.class.pred)
+rf.class.train_err
+rf.class.test_err
+```
+Once the Random Forest tree with two variables was built, the final test MSE was calculated.
+This model produced a test MSE of 0.197, which is significantly lower than the test MSE for the classification model (0.3).
+The Random Forest model was selected as the best method for predicting if a wine were to fall into the "high" or "low" class when it comes to wine quality.
+
+# Regression Methods
+## Multiple Linear Regression
+NEED CODE FROM YUN
+```{r mlr}
+#### REGRESSION ####
+lm.reg <- lm(form.reg, data = trn)
+
+# Obtain RMSE and results for linear model
+lm.reg.pred <- predict(lm.reg, newdata = tst)
+lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
+```
+Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
+The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
+Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
+The formula for a multiple linear regression model with $p$ predictor variables is as follows:
+$$
+\begin{aligned}
+&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
+& \quad \text{where for } i = n \text{ observations}, \\
+& \quad y_i = \text{dependent variable (response)} \\
+& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
+& \quad \beta_0 = \text{y-intercept} \\
+& \quad \beta_j = \text{slope coefficients for each variable} \\
+& \quad \epsilon_i = \text{residuals (error term of model)} \\
+\end{aligned}
+$$
+
+The multiple linear regression model is subject to the following assumptions about the data:
+
+  * Linear relationship between response and predictors
+  * No collinearity between predictor variables
+  * $y_i$'s are iid
+  * $\epsilon_i \sim N(0, \sigma^2)$
+
+To assess these assumptions, we looked at various diagnostics plots:
+```{r lm.diag, out.height = '3.5in'}
+# Look at residual plot for nonlinearity
+regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
+                 aes(x = lm.reg$residuals, y = lm.reg$fitted)) + 
+            geom_point(position = 'jitter') + labs(x = "Residuals", y = "Fitted Values")
+
+# Look at standardized residual plot for outliers
+stdres <- ggplot(data = data.frame(x = 1:length(lm.reg$residuals), 
+                                   y = scale(lm.reg$residuals)), 
+                 aes(y = y, x = x)) + 
+            geom_point() + labs(x = "Index", y = "Standardized Residuals")
+
+# Look at a qqplot to assess normality
+ggqq <- ggplot(data.frame(lm.reg$residuals), aes(sample = lm.reg$residuals)) + 
+          stat_qq() + stat_qq_line(col = "blue", lty = 2) + 
+          labs(x = "Theoretical", y = "Sample")
+
+# Look at residuals vs leverage plot
+reslev <- ggplot(lm.reg, aes(x = .hat, y = .stdresid)) + 
+            geom_point(aes(size = .cooksd)) + 
+            stat_smooth(method = "lm") + 
+            labs( x = "Leverage", y = "Standardized Residuals") + 
+            scale_size_continuous("Cook's D", range = c(1,5)) +
+            theme_bw() + theme(legend.position = "right")
+
+# Stitch all the graphs together
+grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)
+```
+
+~ talk about residual plots here ~
+
+The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
+The coefficients can be computed efficiently and they are have clear and easy interpretations.
+It is also possible to make inference on any of the variables or linear combination of variables.
+The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
+
+## Random Forest for Regression
+As mentioned earlier in the classification section, Random Forest is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees.
+This model will be built exactly in the same way as the classification model, except the tree will not be predicting if the wine is a part of the good or bad class, but rather predicts the quality score associated with each wine.
+As before the model will first be built using cross validation to determine the number of variables that should be used in order to produce the smallest test MSE.
+Similar to the random forest classification tree, the cross validation selects the tree with 2 variables.
+```{r rf regression}
+#----- regression -----
+test.err=double(11)
+for (mtry in 1:11){
+  fit<- randomForest(quality~.-class,trn,mtry=mtry)
+  pred <- predict(fit,tst)
+  test.err[mtry] <- sqrt(mean((tst$quality - pred)^2))
+}
+matplot(1:mtry,test.err,pch=19,col='red',type='b',
+        ylab='Mean Square Error')
+
+rf.reg<-randomForest(quality~.-class,trn,mtry=2)
+rf.reg.pred<-predict(rf.reg,tst)
+rf.reg.train_err<-sqrt(mean((trn$quality-predict(rf.reg))^2))
+rf.reg.test_err<-sqrt(mean((tst$quality-rf.reg.pred)^2))
+rf.reg.train_err
+rf.reg.test_err
+```  
+The test MSE for the Regression Random Forest using 2 variables is 0.55. COMPARE THIS TO MLR WHEN WE GET YUN'S CODE.
+
+I'm not sure we need these next 2 sections because most of the methods we are using do model selection.
+I don't have Yun's code, but I'm assuming he does some sort of variable selection, so this part seems extraneous.
 ## Cross Validation
@@
 We split the data into a training and test data set in order to later evaluate the model's prediction accuracy or root mean squared error (RMSE).
-For this project we used a 75/25 split, training the data on the 75% and testing the trained models on the withheld 25%.
+For this project we used a 80/20 split, training the data on the 80% and testing the trained models on the withheld 20%.
 We will then repeat this process over 5 folds of the data, averaging the results.
@@
 
-## Multiple Linear Regression
-```{r mlr}
-#### REGRESSION ####
-lm.reg <- lm(form.reg, data = trn)
-
-# Obtain RMSE and results for linear model
-lm.reg.pred <- predict(lm.reg, newdata = tst)
-lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
-```
-Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
-The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
-Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
-The formula for a multiple linear regression model with $p$ predictor variables is as follows:
-$$
-\begin{aligned}
-&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
-& \quad \text{where for } i = n \text{ observations}, \\
-& \quad y_i = \text{dependent variable (response)} \\
-& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
-& \quad \beta_0 = \text{y-intercept} \\
-& \quad \beta_j = \text{slope coefficients for each variable} \\
-& \quad \epsilon_i = \text{residuals (error term of model)} \\
-\end{aligned}
-$$
-
-The multiple linear regression model is subject to the following assumptions about the data:
-
-  * Linear relationship between response and predictors
-  * No collinearity between predictor variables
-  * $y_i$'s are iid
-  * $\epsilon_i \sim N(0, \sigma^2)$
-
-To assess these assumptions, we looked at various diagnostics plots:
-```{r lm.diag, out.height = '3.5in'}
-# Look at residual plot for nonlinearity
-regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
-                 aes(x = lm.reg$residuals, y = lm.reg$fitted)) + 
-            geom_point(position = 'jitter') + labs(x = "Residuals", y = "Fitted Values")
-
-# Look at standardized residual plot for outliers
-stdres <- ggplot(data = data.frame(x = 1:length(lm.reg$residuals), 
-                                   y = scale(lm.reg$residuals)), 
-                 aes(y = y, x = x)) + 
-            geom_point() + labs(x = "Index", y = "Standardized Residuals")
-
-# Look at a qqplot to assess normality
-ggqq <- ggplot(data.frame(lm.reg$residuals), aes(sample = lm.reg$residuals)) + 
-          stat_qq() + stat_qq_line(col = "blue", lty = 2) + 
-          labs(x = "Theoretical", y = "Sample")
-
-# Look at residuals vs leverage plot
-reslev <- ggplot(lm.reg, aes(x = .hat, y = .stdresid)) + 
-            geom_point(aes(size = .cooksd)) + 
-            stat_smooth(method = "lm") + 
-            labs( x = "Leverage", y = "Standardized Residuals") + 
-            scale_size_continuous("Cook's D", range = c(1,5)) +
-            theme_bw() + theme(legend.position = "right")
-
-# Stitch all the graphs together
-grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)
-```
-
-~ talk about residual plots here ~
-
-The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
-The coefficients can be computed efficiently and they are have clear and easy interpretations.
-It is also possible to make inference on any of the variables or linear combination of variables.
-The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
-
-## Classification Tree  
-The basic classification tree is based on partitioning the data into subgroups using simple binary splitting.
-Initially, all objects are considered a single group.
-Then, the group is repeatedly binarily split into two subgroups based on the criteria of a certain variable.
-In the classification setting, we classify the observations in a specific region with majority vote.
-We want to grow the tree as big as we can, and then prune it back using cost-complexity pruning.
-This is done to not overfit the data, but pruning increases the bias.
-The pruning parameter needs to be tuned; below is a plot of the classification accuracy by pruning parameter, where .001 yields the highest accuracy.
-
-```{r tree, out.height='3in'}
-# Set seed for reprodicibility
-set.seed(702)
-
-# Fit classification tree model
-tree.class <- train(form.class, data = trn, method = "rpart",
-                    trControl = trctrl, tuneGrid = expand.grid(cp = seq(0, .02, .001)))
-
-# Plot the accuracy as a function of the tuning parameter
-plot(Accuracy ~ cp, data = tree.class$results, xlab = "Cp")
-points(which.max(tree.class$results$Accuracy),
-       tree.class$results$Accuracy[which.max(tree.class$results$Accuracy)],
-       pch = 8, col = "#CC0000")
-```
-
-The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
-Additionally, the classification tree is not very computationally expensive, unlike the random forest model.
-We used the `caret` and `rpart` packages in R to fit our trees.
-
-A visualization of the classification tree can be found below.
-
-```{r, out.heigh = '3.5in'}
-# Obtain confusion matrix and results for H2
-tree.class.pred <- predict(tree.class, newdata = tst)
-tree.class.CM <- confusionMatrix(tree.class.pred, tst$class)
-tree.class.res <- c(tree.class.CM$overall[1], tree.class.CM$byClass[1:2])
-
-# Print out classification tree visualization
-tree.class.out <- rpart(form.class, data = trn, cp = 0.011)
-rpart.plot(tree.class.out, main="Classification Tree")
-```
-
-## Random Forest  
-The random forest model is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results.
-Many bootstrap samples are created with replacement and then a classification tree is fitted on each bootstrap sample with a random subset of the predictors.
-Once a prediction has been made by all of the bootstrapped trees, the final classification is based on majority vote of the bootstrap predictions.
-Similarly, the prediction in the regression context is an average of all of the predictions of the bootstrapped trees.
-
-The following parameters need to be tuned in a random forest model:  
-  * `mtry`: the number of randomly-selected variables selected at a node split
-  * `ntree`: the number of trees to grow
-
-```{r rf, out.height='3.5in'}
-#### REGRESSION ####
-# Set seed for reproducibility
-set.seed(702)
-
-# Fit the regression random forest model 
-rf.reg <- train(form.reg, data = trn, method = "rf",
-                trControl = trctrl, importance = TRUE,
-                tuneGrid = expand.grid(mtry = 1:6))
-
-# Obtain RMSE and results
-rf.reg.pred <- predict(rf.reg, newdata = tst)
-rf.reg.rmse <- sqrt(mean((rf.reg.pred - tst$quality)^2))
-
-#### CLASSIFICATION ####
-# Fit the classification random forest model
-rf.class <- train(form.class, data = trn, method = "rf",
-                trControl = trctrl, importance = TRUE,
-                tuneGrid = expand.grid(mtry = 1:5))
-
-# Obtain confusion matrix and results
-rf.class.pred <- predict(rf.class, newdata = tst)
-rf.class.CM <- confusionMatrix(rf.class.pred, tst$class)
-rf.class.res <- c(rf.class.CM$overall[1], rf.class.CM$byClass[1:2])
-
-# Plot the RMSE and accuracy as a function of mtry (ntree constant at 500)
-par(mfrow = c(1,2))
-plot(RMSE ~ mtry, data = rf.reg$results)
-points(which.min(rf.reg$results$RMSE), rf.reg$results$RMSE[which.min(rf.reg$results$RMSE)],
-       pch = 8, col = "#CC0000")
-plot(Accuracy ~ mtry, data = rf.class$results)
-points(which.max(rf.class$results$Accuracy),
-       rf.class$results$Accuracy[which.max(rf.class$results$Accuracy)],
-       pch = 8, col = "#CC0000")
-```
-
-Since `ntree` will plateau with enough trees, we tuned the `mtry` in both the regression and classification settings.
-For regression, the optimal `mtry` value is `r rf.reg$bestTune[1,1]`, while for classification it is  `r rf.class$bestTune[1,1]`.
-Keep in mind that the random forest is equivalent to bagging when the number of randomly selected features is equal to the total number of variables in the model.
-In both of the above plots, this is represented by the maximum displayed value of `mtry`.
-
-The advantage for using an ensemble method over a regular classification or regression tree is that because there are many bootstrap samples being averaged together, there is less variance.
-This is similar to how the variance of the sample mean goes down as the sample size increases.
-Although it will probably increase our prediction accuracy, the random forest loses the interpretability that the basic trees have.
-Furthermore, the algorithm that is used to fit the random forest is very computationally expensive.
-
 # Results
@@
 # Create data frame output for classification models
-df.class <- rbind.data.frame(tree.class.res, rf.class.res) %>% round(4)
-df.class$method <- cl
-colnames(df.class) <- c("A", "B", "C", "D")
-df.class <- df.class %>% select("D", "A", "B", "C")
-colnames(df.class) <- c("Classification Method", "Accuracy", "Sensitivity", "Specificity")
+df.class <- rbind.data.frame(prune.tree.class.test_err, rf.class.test_err) %>% round(4)
+colnames(df.class) <- c("Misclassification Rate")
+rownames(df.class) <- c("Classification Model", "Random Forest")
+df.class
 ```
@@
 As an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
-We suspect that other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracies.
+We suspect that other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracy.
 

### 2021-04-20 DRAFT.Rmd
mana edits

@@
 bad.wine <- subset(wine, quality <= 5)
-summary(good.wine)
-summary(bad.wine)
+# summary(good.wine)
+# summary(bad.wine)
 ```
@@
 # 1. full size decision tree
-tree.class<-tree(class~.-quality,trn)
-plot(tree.class); text(tree.class, pretty=0)
-tree.class.pred<-predict(tree.class,tst,type='class')
-tree.class.train_err<-mean(trn$class!=predict(tree.class,type='class'))
-tree.class.test_err<-mean(tst$class!=tree.class.pred)
-tree.class.train_err
-tree.class.test_err
+tree.class <- tree(class ~ . - quality, trn)
+plot(tree.class); text(tree.class, pretty = 0)
+tree.class.pred <- predict(tree.class, tst, type = 'class')
+tree.class.train_err <- mean(trn$class != predict(tree.class, type = 'class'))
+tree.class.test_err <- mean(tst$class != tree.class.pred)
+# tree.class.train_err
+# tree.class.test_err
 ```
-The resulting classification tree creates splits on 4 variables, Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity.
-There are 10 terminal nodes.
-The tree has a test error rate of 0.297.
+
+The resulting classification tree creates splits on 4 variables: Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity.
+The tree has a test error rate of `r round(tree.class.test_err, 3)`.
 This test rate is fairly low, but as mentioned before, if one conducts pruning on the tree, the error rate can decrease and a more accurate model is created.
@@
 The resulting pruned model will now have 8 terminal nodes and still uses the four variables in the un-pruned tree.
+
 ```{r pruned tree}
 # 2. pruned decision tree 
-cv.tree.class<-cv.tree(tree.class,FUN=prune.misclass)
+cv.tree.class <- cv.tree(tree.class, FUN = prune.misclass)
 plot(cv.tree.class)
-prune.tree.class<-prune.misclass(tree.class,best=8)
-plot(prune.tree.class);text(prune.tree.class,pretty=0)
-prune.tree.class.pred<-predict(prune.tree.class,tst,type='class')
-prune.tree.class.train_err<-mean(trn$class!=predict(prune.tree.class,type='class'))
-prune.tree.class.test_err<-mean(tst$class!=prune.tree.class.pred)
-prune.tree.class.test_err
+prune.tree.class <- prune.misclass(tree.class, best = 8)
+plot(prune.tree.class); text(prune.tree.class, pretty = 0)
+prune.tree.class.pred <- predict(prune.tree.class, tst, type = 'class')
+prune.tree.class.train_err <- mean(trn$class != predict(prune.tree.class, type = 'class'))
+prune.tree.class.test_err <- mean(tst$class != prune.tree.class.pred)
+# prune.tree.class.test_err
 ```
-The pruned tree produces a test error rate to the original tree with a test MSE of 0.3.
-The pruned tree would be selected using the concept of Occam's Razor, where the simplest model is the best model choice.
 
-Using a Random Forest tree can often improve upon the error rate of a simple classification tree.
-While the test MSE was small in the classification tree, the team believed the error rate could be decreases in using a more complex tree based method.
+The pruned tree produces a test error rate to the original tree with a misclassification error of `r prune.tree.class.test_err`.Using a Random Forest tree can often improve upon the error rate of a simple classification tree.
+While the test MSE was small in the classification tree, the team believed the error rate could be decreased using a more complex tree based method.
 
@@
 A plot of these is included below, the tree that produced the smallest error rate was the tree built with two variables, so that was the tree selected through cross validation.
+
 ```{r rf classification CV}
@@
 for (mtry in 1:11){
-  fit <- randomForest(class~.-quality, trn,mtry=mtry)
-  pred <- predict(fit,tst)
+  fit <- randomForest(class ~ . - quality, trn, mtry = mtry)
+  pred <- predict(fit, tst)
   test.err[mtry] <- mean(tst$class != pred)
 }
-matplot(1:mtry,test.err,pch=19,col='red',type='b',
-        ylab='Mean Square Error')
+
+# Plot cross validation results
+matplot(1:mtry, test.err, pch = 1, type = 'b',
+        xlab = "mtry", ylab = 'Mean Square Error')
+points(which.min(test.err), test.err[which.min(test.err)], pch = 8, col = "#CC0000")
 ```
@@
 ```{r rf class models}
-rf.class <- randomForest(class~.-quality,trn,mtry=2)
+rf.class <- randomForest(class ~ . - quality, trn, mtry = 2)
 rf.class.pred <- predict(rf.class,tst)
-rf.class.train_err <- mean(trn$class!=predict(rf.class))
-rf.class.test_err <- mean(tst$class!=rf.class.pred)
-rf.class.train_err
-rf.class.test_err
+rf.class.train_err <- mean(trn$class != predict(rf.class))
+rf.class.test_err <- mean(tst$class != rf.class.pred)
+# rf.class.train_err
+# rf.class.test_err
 ```
+
 Once the Random Forest tree with two variables was built, the final test MSE was calculated.
-This model produced a test MSE of 0.197, which is significantly lower than the test MSE for the classification model (0.3).
-The Random Forest model was selected as the best method for predicting if a wine were to fall into the "high" or "low" class when it comes to wine quality.
+This model produced a test MSE of `r round(rf.class.test_err, 3)`, which is significantly lower than the test MSE for the basic classification tree (`r round(prune.tree.class.test_err`)).
+The random forest was selected as the best method for predicting if a wine were to fall into the "high" or "low" class when it comes to wine quality.
 
@@
 #### REGRESSION ####
-lm.reg <- lm(form.reg, data = trn)
+lm.reg <- lm(quality ~ . - class, data = trn)
 
@@
 Similar to the random forest classification tree, the cross validation selects the tree with 2 variables.
+
 ```{r rf regression}
 #----- regression -----
-test.err=double(11)
+# loop thru all the different values of mtry
+test.err = double(11)
 for (mtry in 1:11){
-  fit<- randomForest(quality~.-class,trn,mtry=mtry)
-  pred <- predict(fit,tst)
+  fit<- randomForest(quality ~ . - class, trn, mtry = mtry)
+  pred <- predict(fit, tst)
   test.err[mtry] <- sqrt(mean((tst$quality - pred)^2))
 }
-matplot(1:mtry,test.err,pch=19,col='red',type='b',
-        ylab='Mean Square Error')
 
-rf.reg<-randomForest(quality~.-class,trn,mtry=2)
-rf.reg.pred<-predict(rf.reg,tst)
-rf.reg.train_err<-sqrt(mean((trn$quality-predict(rf.reg))^2))
-rf.reg.test_err<-sqrt(mean((tst$quality-rf.reg.pred)^2))
-rf.reg.train_err
-rf.reg.test_err
+# Plot cross validation results
+matplot(1:mtry, test.err, pch = 1, type = 'b', ylab = 'Mean Square Error', xlab = "mtry")
+points(which.min(test.err), test.err[which.min(test.err)], pch = 8, col = "#CC0000")
+
+# fit the random forest with the best results
+rf.reg <- randomForest(quality ~ . - class, trn, mtry = 2)
+rf.reg.pred <- predict(rf.reg, tst)
+rf.reg.train_err <- sqrt(mean((trn$quality - predict(rf.reg))^2))
+rf.reg.test_err <- sqrt(mean((tst$quality - rf.reg.pred)^2))
+#rf.reg.train_err
+#rf.reg.test_err
 ```  
 The test MSE for the Regression Random Forest using 2 variables is 0.55. COMPARE THIS TO MLR WHEN WE GET YUN'S CODE.
-
-I'm not sure we need these next 2 sections because most of the methods we are using do model selection.
-I don't have Yun's code, but I'm assuming he does some sort of variable selection, so this part seems extraneous.
-## Cross Validation
-Again, our focus is on the classification accuracy of machine learning methods.
-In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the models.
-We split the data into a training and test data set in order to later evaluate the model's prediction accuracy or root mean squared error (RMSE).
-For this project we used a 80/20 split, training the data on the 80% and testing the trained models on the withheld 20%.
-We will then repeat this process over 5 folds of the data, averaging the results.
-
-```{r CV}
-# Specify the cross-validation method (5-fold CV)
-trctrl <- trainControl(method = "cv", number = 5)
-```
-
-## Variable Selection
-Since we have two different tasks that we are trying to complete (classification and regression), we went about selecting the variables for each task differently.
-For the classification task, we ran an initial random forest model on the entire dataset with selected predictors.
-We then looked at the variable importance of the predictors to decide which variables we would use in our final models.
-Within the context of machine learning, variable importance refers to how much a given model "uses" that variable to make accurate predictions.
-In other words, the more a model relies on a variable to make predictions, the more important it is for the model.
-
-```{r class.select, out.height='4in'}
-# Write preliminary model  
-model.class.pre <- class ~ .
-
-# Fit preliminary random forest on entire data set
-rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
-                             ntree = 500, mtry = 3, importance = TRUE)
-
-# Create data frame of variable importance
-df.class.pre <- rf.class.pre$importance %>% as.data.frame()
-X <- rownames(df.class.pre)
-
-# Sort the data frame and prep for plot
-VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
-VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
-VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)
-
-# Plot variable importance
-vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
-                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
-                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
-                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
-                ggtitle("Variable Importance")
-vimp.class
-```
-
-We decided to choose the five variables with the most variable importance, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`.
-The fact that density proved to be an important variable for classification was relatively surprising.
-One would think that the wine quality would be determined based on its attributes that are detectable by the human senses.
-For this same reason, alcohol, acidity, and sulphates made sense in terms of variable importance.
-Sulphates are essential for the preservation of the wine, an overly- or underly-acidic wine may be considered poor, and the alcohol content has a connection to the fermentation process, where the must (aka the grape juice before it undergoes fermentation) actually becomes wine.
-
-When looking at the problem in terms of a regression context, we used best subsets selection for feature selection.
-Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors.
-Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
-We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
-
-```{r reg.select, out.height='3.5in'}
-# Create model and get summary output
-lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
-smre <- lm.sub %>% summary()
-
-# Plot the Cp, BIC, R^2, RSS results in one plot
-par(mfrow = c(2,2))
-plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
-points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
-plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
-points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
-plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
-points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
-plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
-points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
-
-# Best subset selection
-best.betas <- coef(lm.sub, id = 6) %>% round(2)
-```
-
-Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
-Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
-This was the model favored by BIC, which gives the greatest penalty for complexity in models.
-The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
-All of the variables from the classification selection appear here with the exception of density.
-In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses.
-The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.
-
-```{r form}
-# Write formula for regression models
-form.reg <- quality ~ volatile.acidity + chlorides + pH + 
-            total.sulfur.dioxide + sulphates + alcohol
-
-# Write formula for classification models
-form.class <- class ~ alcohol + sulphates + density +
-              volatile.acidity + total.sulfur.dioxide
-```
 
@@
 # Create data frame output for regression models
-df.reg <- rbind.data.frame(lm.reg.rmse, rf.reg.rmse) %>% round(4)
+df.reg <- rbind.data.frame(lm.reg.rmse, rf.reg.test_err) %>% round(4)
 df.reg$method <- reg
@@
 colnames(df.class) <- c("Misclassification Rate")
-rownames(df.class) <- c("Classification Model", "Random Forest")
-df.class
+rownames(df.class) <- c("Classification Tree", "Random Forest")
 ```
@@
 It would be interesting to see how different weather conditions, terroirs, plant ages, and slopes affect the outcome of the wine in terms of quality.
+
+# NOT NEEDED ???
+
+I'm not sure we need these next 2 sections because most of the methods we are using do model selection.
+I don't have Yun's code, but I'm assuming he does some sort of variable selection, so this part seems extraneous.
+
+## Cross Validation
+Again, our focus is on the classification accuracy of machine learning methods.
+In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the models.
+We split the data into a training and test data set in order to later evaluate the model's prediction accuracy or root mean squared error (RMSE).
+For this project we used a 80/20 split, training the data on the 80% and testing the trained models on the withheld 20%.
+We will then repeat this process over 5 folds of the data, averaging the results.
+
+```{r CV, eval = F}
+# Specify the cross-validation method (5-fold CV)
+trctrl <- trainControl(method = "cv", number = 5)
+```
+
+## Variable Selection
+Since we have two different tasks that we are trying to complete (classification and regression), we went about selecting the variables for each task differently.
+For the classification task, we ran an initial random forest model on the entire dataset with selected predictors.
+We then looked at the variable importance of the predictors to decide which variables we would use in our final models.
+Within the context of machine learning, variable importance refers to how much a given model "uses" that variable to make accurate predictions.
+In other words, the more a model relies on a variable to make predictions, the more important it is for the model.
+
+```{r class.select, out.height='4in', eval = F}
+# Write preliminary model  
+model.class.pre <- class ~ .
+
+# Fit preliminary random forest on entire data set
+rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
+                             ntree = 500, mtry = 3, importance = TRUE)
+
+# Create data frame of variable importance
+df.class.pre <- rf.class.pre$importance %>% as.data.frame()
+X <- rownames(df.class.pre)
+
+# Sort the data frame and prep for plot
+VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
+VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
+VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)
+
+# Plot variable importance
+vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
+                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
+                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
+                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
+                ggtitle("Variable Importance")
+vimp.class
+```
+
+We decided to choose the five variables with the most variable importance, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`.
+The fact that density proved to be an important variable for classification was relatively surprising.
+One would think that the wine quality would be determined based on its attributes that are detectable by the human senses.
+For this same reason, alcohol, acidity, and sulphates made sense in terms of variable importance.
+Sulphates are essential for the preservation of the wine, an overly- or underly-acidic wine may be considered poor, and the alcohol content has a connection to the fermentation process, where the must (aka the grape juice before it undergoes fermentation) actually becomes wine.
+
+When looking at the problem in terms of a regression context, we used best subsets selection for feature selection.
+Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors.
+Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
+We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
+
+```{r reg.select, out.height='3.5in', eval = F}
+# Create model and get summary output
+lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
+smre <- lm.sub %>% summary()
+
+# Plot the Cp, BIC, R^2, RSS results in one plot
+par(mfrow = c(2,2))
+plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
+points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
+plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
+points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
+plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
+points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
+plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
+points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
+
+# Best subset selection
+best.betas <- coef(lm.sub, id = 6) %>% round(2)
+```
+
+Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
+Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
+This was the model favored by BIC, which gives the greatest penalty for complexity in models.
+The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
+All of the variables from the classification selection appear here with the exception of density.
+In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses.
+The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.
+
+```{r form, eval = F}
+# Write formula for regression models
+form.reg <- quality ~ volatile.acidity + chlorides + pH + 
+            total.sulfur.dioxide + sulphates + alcohol
+
+# Write formula for classification models
+form.class <- class ~ alcohol + sulphates + density +
+              volatile.acidity + total.sulfur.dioxide
+```

### 2021-04-20 DRAFT.Rmd
mana edit 4/19/2021 see email for more details of edit

@@
 knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
-                      warning = FALSE, fig.align = "center")
+                      warning = FALSE, fig.align = "center",
+                      out.height = '3.5in')
 
@@
 library(rmarkdown) # for output documents
-library(rpart) # fitting classification tree
-library(rpart.plot) # visualization of classification tree
 library(randomForest) # fitting random forest model
@@
 library(gridExtra) # stitching graphs together
-library(tree)
+library(tree) # fitting the classification tree
 
@@
   * `knitr`: for rmarkdown table outputs
+  * `kableExtra`: for making fancy tables
   * `rmarkdown`: for output documents
   * `corrplot`: for correlation structure visualization
-  * `leaps`: for subset selection
-  * `rpart`: fitting the basic classification tree
-  * `rpart.plot`: visualization of the classification tree
+  * `tree`: fitting the basic classification tree
   * `randomForest`: fitting random forest models
+  * `gridExtra`: for stitching plots together
 
@@
 Each entry in the dataset represents the different metrics of the following attributes of a single type of wine.
+
   * **fixed acidity**: the quantity of fixed acids found in the wines.
@@
 ## Data Exploration
-The first thing that was done to the data upon reading it was to use the numeric variables and create pairwise scatterplots.
+The first thing that was done to the data upon reading it was to look at histograms of all of the variables. In doing so, we cann get a feel for the distributions of the individual variables.
+
+```{r}
+# Create histograms for all 12 variables
+par(mfrow = c(3,4))
+for (i in 1:ncol(wine)){
+  hist(wine[,i], xlab = names(wine)[i], 
+       main = names(wine)[i])
+}
+```
+
+Most wines are moderate wines, with less excellent amd poor wines.
+The density and pH of the wine seem to be normally distributed, while the sulfur, acid, and sulphate content seem skewed to the right.
+The density varies very little; looking at the scale on the distribution, we can see that it is approximately normally distributed around .997 with a standard deviation of about .002.
+The distribution of alcohol content came as no surprise; an average bottle of wine will range from about 10-12% alcohol by volume, with some exceptions.
+Beyond 16%, the wine is then considered fortified wine, and red wines don't usually have alcohol content below about 8%.
+
+Next, we went ahead and looked at pairwise scatterplots of the variables.
 This way, one can obtain a greater understanding of how the variables relate to one another, as well as if there is any correlation structure and what it looks like.
@@
 
-```{r pairs, out.width='4in', out.height='3.5in'}
+```{r pairs}
 # Create pairwise scatterplots
@@
 
-```{r corrplot, echo = FALSE, out.height='3.5in'}
+```{r corrplot}
 # Create visualization of correlation structure
@@
 
-After a split in the data was performed to pull out the wines classified as good and the wines classified as bad.
-Further analysis was performed on each group to see if any of the variables were significantly different between the two groups.
+After a split in the data was performed to pull out the wines classified as good and the wines classified as bad, further analysis was performed on each group to see if any of the variables were significantly different between the two groups.
 
-```{r good v bad}
+```{r goodvbad}
 good.wine <- subset(wine, quality > 5)
@@
 # summary(bad.wine)
+
+# Express summary output in table form
+good.out <- array(rep(0, 12*6), c(12,6))
+bad.out <- array(rep(0, 12*6), c(12,6))
+for(i in 1:11){
+  good.out[i,] <- summary(good.wine[,i]) %>% as.vector()
+  bad.out[i,] <- summary(bad.wine[,i]) %>% as.vector()
+}
+rownames(good.out) <- rownames(bad.out) <- names(good.wine)
+colnames(good.out) <- colnames(bad.out) <- c("Min.", "Q1", "Median", "Mean", "Q3", "Max.")
+
+kable(good.out, caption = "Summary Output for Good Wines") %>%
+  kable_styling(latex_options = "HOLD_position")
+kable(bad.out, caption = "Summary Output for Bad Wines") %>%
+  kable_styling(latex_options = "HOLD_position")
 ```
+
 The two groups were similar in most of the variables.
@@
 
-# Create train/test data for H1
+# Create train/test data
 trn <- wine[ind,]
@@
 ## Classification Tree
-
-THIS SECTION NEEDS TO BE EDITED
-
 The basic classification tree is based on partitioning the data into subgroups using simple binary splitting.
@@
 In the classification setting, an observation is classified in a specific region with majority vote.
-Trees are grown as big as possible and then pruned using cost-complexity pruning.
+Trees are grown as big as possible and then pruned back using cost-complexity pruning.
 This is done to ensure the data is not being overfit.
 The tradeoff here is that pruning increases the bias.
-The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
-Additionally, the classification tree is not very computationally expensive, unlike the random forest model.
 
-```{r tree, out.height='3in'}
+```{r tree}
 # Set seed for reprodicibility
@@
 tree.class <- tree(class ~ . - quality, trn)
-plot(tree.class); text(tree.class, pretty = 0)
+# plot(tree.class); text(tree.class, pretty = 0)
 tree.class.pred <- predict(tree.class, tst, type = 'class')
@@
 
-The resulting classification tree creates splits on 4 variables: Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity.
-The tree has a test error rate of `r round(tree.class.test_err, 3)`.
-This test rate is fairly low, but as mentioned before, if one conducts pruning on the tree, the error rate can decrease and a more accurate model is created.
+Pruning is performed on the tree using cross validation.
+A tree is built many times with a different number of splits and the tree that produces the smallest error rate is selected.
+Cross validation on this model shows that 8 is the optimal number of splits to use.
+The resulting pruned model will now have 8 terminal nodes and uses four variables: Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity.
 
-Pruning is performed on the tree using cross validation, the tree is built many times with a different number of splits and the tree that produces the smallest error rate is selected.
-Cross validation on this model shows that the best model is the model built with 8 splits.
-The resulting pruned model will now have 8 terminal nodes and still uses the four variables in the un-pruned tree.
-
-```{r pruned tree}
+```{r prunedtree, out.height = '4in'}
 # 2. pruned decision tree 
+par(mfrow = c(1,2))
 cv.tree.class <- cv.tree(tree.class, FUN = prune.misclass)
@@
 prune.tree.class <- prune.misclass(tree.class, best = 8)
-plot(prune.tree.class); text(prune.tree.class, pretty = 0)
+plot(prune.tree.class); text(prune.tree.class, pretty = 0, cex = .6)
 prune.tree.class.pred <- predict(prune.tree.class, tst, type = 'class')
@@
 prune.tree.class.test_err <- mean(tst$class != prune.tree.class.pred)
+tree.res <- c(prune.tree.class.train_err, prune.tree.class.test_err)
 # prune.tree.class.test_err
@@
 
-The pruned tree produces a test error rate to the original tree with a misclassification error of `r prune.tree.class.test_err`.Using a Random Forest tree can often improve upon the error rate of a simple classification tree.
-While the test MSE was small in the classification tree, the team believed the error rate could be decreased using a more complex tree based method.
+The pruned tree produces a test error rate to the original tree with a misclassification error of `r prune.tree.class.test_err`.
+The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
+Additionally, the classification tree is not very computationally expensive, unlike ensemble classification methods.
+That being said, using ensemble methods like the random forest can often improve the error rate of a simple classification tree, since it is an average of many classification trees.
+While the test MSE was small in the classification tree, we decided to investigate the performance of a random forest, hoping to improve the accuracy of our predictions.
 
-## Random Forest Classification
+## Random Forest for Classification
 The random forest model is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees.
@@
 
-The first thing that should be done in building a boosted tree is cross validation.
+The random forest model requires tuning of the number of randomly selected features at each split using cross validation.
 Here the model is built many times with the number of predictors included from one to all predictors.
 The test MSE is evaluated for each and the smallest value is selected.
-A plot of these is included below, the tree that produced the smallest error rate was the tree built with two variables, so that was the tree selected through cross validation.
+A plot of mtry as a function of the classification error is included below.
 
@@
 matplot(1:mtry, test.err, pch = 1, type = 'b',
-        xlab = "mtry", ylab = 'Mean Square Error')
+        xlab = "mtry", ylab = 'Misclassification Error')
 points(which.min(test.err), test.err[which.min(test.err)], pch = 8, col = "#CC0000")
@@
 
-```{r rf class models}
+```{r rfclassmodels}
 rf.class <- randomForest(class ~ . - quality, trn, mtry = 2)
@@
 rf.class.test_err <- mean(tst$class != rf.class.pred)
+rf.class.res <- c(rf.class.train_err, rf.class.test_err)
 # rf.class.train_err
@@
 
-Once the Random Forest tree with two variables was built, the final test MSE was calculated.
-This model produced a test MSE of `r round(rf.class.test_err, 3)`, which is significantly lower than the test MSE for the basic classification tree (`r round(prune.tree.class.test_err`)).
-The random forest was selected as the best method for predicting if a wine were to fall into the "high" or "low" class when it comes to wine quality.
+From cross validation, the model that produced the smallest error rate was the model that was built with   `r which.min(test.err)` randomly-chosen variables at each split.
+This model produced a test error of `r round(rf.class.test_err, 3)`, which is significantly lower than the test error for the basic classification tree (`r round(prune.tree.class.test_err, 3)`).
+The random forest was selected as the best method for predicting if a wine were to fall into the "high" or "low" quality desgination.
 
@@
 # Obtain RMSE and results for linear model
+lm.reg.trn.pred <- predict(lm.reg, newdata = trn)
+lm.reg.trn.rmse <- sqrt(mean((lm.reg.trn.pred - tst$quality)^2))
+
 lm.reg.pred <- predict(lm.reg, newdata = tst)
 lm.reg.rmse <- sqrt(mean((lm.reg.pred - tst$quality)^2))
+lm.res <- c(lm.reg.trn.rmse, lm.reg.rmse)
 ```
+
 Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
@@
 The formula for a multiple linear regression model with $p$ predictor variables is as follows:
+
 $$
@@
 To assess these assumptions, we looked at various diagnostics plots:
-```{r lm.diag, out.height = '3.5in'}
+```{r lm.diag}
 # Look at residual plot for nonlinearity
@@
 
-```{r rf regression}
+```{r rf.reg}
 #----- regression -----
@@
 # Plot cross validation results
-matplot(1:mtry, test.err, pch = 1, type = 'b', ylab = 'Mean Square Error', xlab = "mtry")
+matplot(1:mtry, test.err, pch = 1, type = 'b', ylab = 'Root MSE', xlab = "mtry")
 points(which.min(test.err), test.err[which.min(test.err)], pch = 8, col = "#CC0000")
@@
 rf.reg.test_err <- sqrt(mean((tst$quality - rf.reg.pred)^2))
-#rf.reg.train_err
-#rf.reg.test_err
+rf.reg.res <- c(rf.reg.train_err, rf.reg.test_err)
 ```  
@@
 # Create data frame output for regression models
-df.reg <- rbind.data.frame(lm.reg.rmse, rf.reg.test_err) %>% round(4)
-df.reg$method <- reg
-colnames(df.reg) <- c(LETTERS[1:2])
-df.reg <- df.reg %>% select("B", "A")
-colnames(df.reg) <- c("Regression Method", "Root MSE")
+df.reg <- rbind.data.frame(lm.res, rf.reg.res) %>% round(4)
+rownames(df.reg) <- c("Multiple Linear Regression", "Random Forest")
+colnames(df.reg) <- c("Training RMSE", "Test RMSE")
 
 # Create data frame output for classification models
-df.class <- rbind.data.frame(prune.tree.class.test_err, rf.class.test_err) %>% round(4)
-colnames(df.class) <- c("Misclassification Rate")
+df.class <- rbind.data.frame(tree.res, rf.class.res) %>% round(4)
+colnames(df.class) <- c("Training Error", "Test Error")
 rownames(df.class) <- c("Classification Tree", "Random Forest")

### 2021-04-20 DRAFT.Rmd
Added MLR code and descriptions.

@@
 # Set default chunk options
-knitr::opts_chunk$set(echo = FALSE, message = FALSE, 
+knitr::opts_chunk$set(echo = FALSE, message = FALSE, cache = FALSE,
                       warning = FALSE, fig.align = "center",
@@
 ## Multiple Linear Regression
-NEED CODE FROM YUN
+Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
+The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
+Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
+The formula for a multiple linear regression model with $p$ predictor variables is as follows:
+
+$$
+\begin{aligned}
+&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
+& \quad \text{where observation } i = 1, ..., n \\
+& \quad Y_i = \text{dependent variable (response)} \\
+& \quad X_{ij} = \text{independent variables for } j = 1, ... , p \\
+& \quad \beta_0 = \text{y-intercept} \\
+& \quad \beta_j = \text{slope coefficients for each variable} \\
+& \quad \epsilon_i = \text{residuals (error term of model)} \\
+\end{aligned}
+$$
+Although it is possible to use every predictor in the MLR model, we instead use variable selection to reduce the number of predictors to six (`volatile.acidity`, `chlorides`, `pH`, `total.sulfur.dioxide`, `sulphates`, and `alcohol`) while maintaining predictive accuracy (see Variable Selection section for details).
+Applying the MLR with the reduced model, we get the following:
+
 ```{r mlr}
 #### REGRESSION ####
-lm.reg <- lm(quality ~ . - class, data = trn)
+lm.reg <- lm(quality ~ volatile.acidity + chlorides + pH + total.sulfur.dioxide + sulphates + alcohol, data = trn)
+summary(lm.reg)
 
@@
 
-Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
-The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
-Like simple linear regression, coefficient estimates are found by minimizing the sum of the squared errors.
-The formula for a multiple linear regression model with $p$ predictor variables is as follows:
-
-$$
-\begin{aligned}
-&Y_i = \beta_0 + \beta_1X_{i1} + \beta_2X_{i2} + ... + \beta_pX_{ip} + \epsilon_i, \\
-& \quad \text{where for } i = n \text{ observations}, \\
-& \quad y_i = \text{dependent variable (response)} \\
-& \quad x_{ji} = \text{indpendent variables } (j = 1, ... , p) \\
-& \quad \beta_0 = \text{y-intercept} \\
-& \quad \beta_j = \text{slope coefficients for each variable} \\
-& \quad \epsilon_i = \text{residuals (error term of model)} \\
-\end{aligned}
-$$
+The adjusted R-squared value 0.3633 suggests that these six predictors combined explain about 36.33% of the variance in `quality` of a wine.
 
@@
   * No collinearity between predictor variables
-  * $y_i$'s are iid
+  * $Y_i$'s are iid
   * $\epsilon_i \sim N(0, \sigma^2)$
@@
 regres <- ggplot(data = data.frame(lm.reg$residuals, lm.reg$fitted.values), 
-                 aes(x = lm.reg$residuals, y = lm.reg$fitted)) + 
-            geom_point(position = 'jitter') + labs(x = "Residuals", y = "Fitted Values")
+                 aes(y = lm.reg$residuals, x = lm.reg$fitted)) + 
+            geom_point(position = 'jitter') + labs(y = "Residuals", x = "Fitted Values")
 
@@
 grid.arrange(regres, stdres, ggqq, reslev, nrow = 2)
+
+scale.res <- scale(lm.reg$residuals)
+scale.res.2std <- mean((scale.res <= 1.96) * (scale.res >= -1.96))
 ```
 
-~ talk about residual plots here ~
+Residuals indicate that our MLR assumptions are mostly satisfied.
+The residuals vs.
+fitted values plot (top left) has unusual stratified lines, but this is caused by the actual responses being integer values of 4, 5, 6, 7, etc.
+rather than continuous values.
+The plot does not show heteroscedasticity, and residuals are spread evenly around 0 over the entire range of fitted values.
+When residuals are standardized, as in the top right plot, we expect about 95% to be inside the [-1.96, 1.96] interval; the actual proportion is `r round(scale.res.2std, 3)`.
+The Q-Q plot at bottom left compares the normality of residuals against a theoretical distribution.
+Except for a slight downward bias of the lowest 5% of residuals at 2 standard deviations below 0, the data fits a normal distribution almost perfectly.
+On the bottom right plot, high leverage points (those with leverage $\ge$ 0.03) are not dramatically biased, being spread both above and below the 0 residual.
+Cook's Distance top out at about 0.05, well below the generally accepted threshold of 1 needed to quality as highly influential.
+These plots taken together suggest that the data generally fulfills the requirements for modeling using MLR.
 
@@
 
-```{r reg.select, out.height='3.5in', eval = F}
+```{r reg.select, out.height='3.5in', eval = T}
 # Create model and get summary output

### 2021-04-20 DRAFT.Rmd
Output MLR model summary to nicer-looking table.

@@
 library(tree) # fitting the classification tree
+library(jtools) # for outputting nice-looking model summaries
 
@@
 lm.reg <- lm(quality ~ volatile.acidity + chlorides + pH + total.sulfur.dioxide + sulphates + alcohol, data = trn)
-summary(lm.reg)
+summ(lm.reg, digits = 3)
 
@@
 
-The adjusted R-squared value 0.3633 suggests that these six predictors combined explain about 36.33% of the variance in `quality` of a wine.
+The adjusted R-squared value 0.363 suggests that these six predictors combined explain about 36.3% of the variance in `quality` of a wine.
 

### 2021-04-20 DRAFT.Rmd
mana final draft edit
moved the reserach question and hypothesis section and gave it its own paragraph.

added a little more to the results section to talk about support of the hypotheses.

Added in the variable selection section using best subsets

changed the required libraries. added in leaps and jtools, got rid of caret #rip

fixed some weird sentences and spelling errors. proofread Yun's addition.

Got rid of the "NOT NECESSARY?" section at the bottom.

good/bad --> low/high

@@
 library(knitr) # making nice tables
-library(caret) # training models & CV
 library(leaps) # subset selection
@@
 Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality.
-The question this study is most interested in is to determine if there is a way to predict the quality of a wine without having to use taste testers.
-The team is hypothesizing that both the classifation and regression methods can accurately describe the quality of wine.
 
@@
 This will be done by performing both regression and classification methods, with regression ranking the quality of the wines on a scale of zero to ten, with zero being the lowest quality and ten being the highest quality.
-The classification models will be classifying the wines into two categories, good and bad, with wines in the bad category having a score of zero to five, and wines in the good category having a quality score between six and ten.
+The classification models will be classifying the wines into two categories, high and low, with wines in the low category having a score of zero to five, and wines in the high category having a quality score between six and ten.
 The assessment of the prediction accuracy will be performed by the following methods:
@@
 
+The research question of interest for this study is whether or not machine learning methods can accurately predict the quality of a wine, independent of using taste testers.
+In the classification setting, we hypothesize that the machine learning methods can correctly classify wines as "high" or "low" quality with 80% accuracy.
+In the context of regression, we hypothesize that the regression techniques used will be able to achieve a low root mean squared error.
+In both settings, though, we expect the random forest to outperform the other methods.
+
 # Required Libraries
@@
 
-  * `caret`: to do the heavy lifting of training and tuning the models
   * `tidyverse`: for all the data reading and wrangling
@@
   * `corrplot`: for correlation structure visualization
+  * `leaps`: for best subsets selection
   * `tree`: fitting the basic classification tree
@@
   * `gridExtra`: for stitching plots together
+  * `jtools`: for linear model summary output
 
@@
 ```
-Based on the pairwise scatterplots, one can conclude that pH may be linearly correlated with density, sulphates, and alcohol.
+
+Based on the pairwise scatterplots, one may conclude that pH may be linearly correlated with density, sulphates, and alcohol.
 This leads us to believe that there is some relationship between the chemical properties of the wines.
@@
 
-After a split in the data was performed to pull out the wines classified as good and the wines classified as bad, further analysis was performed on each group to see if any of the variables were significantly different between the two groups.
+After a split in the data was performed to pull out the wines classified as high quality and the wines classified as low quality, further analysis was performed on each group to see if any of the variables were significantly different between the two groups.
 
@@
 The two groups were similar in most of the variables.
-Two notable differences are that the wines classified as "good" had a lower mean value than "bad" wines for total sulfur dioxide.
-Good wines had a mean of 39, which was significantly lower than the mean of the "bad" wines at 54.6.
+Two notable differences are that the wines classified as high quality had a lower mean value than low quality wines for total sulfur dioxide.
+High quality wines had a mean of 39, which was significantly lower than the mean of the low quality wines at 54.6.
 Another difference to note is the difference in the mean for alcohol.
-The good wines had a mean alcohol content at 10.8, while the bad wines had a lower mean alcohol content around 9.9.
-There were also differences in fixed acidity, with the bad wines having a lower mean.
+The high quality wines had a mean alcohol content at 10.8, while the low quality wines had a lower mean alcohol content around 9.9.
+There were also differences in fixed acidity, with the lower quality wines having a lower mean.
 There were also several variables that don't seem to change between the two groups including pH, residual sugar, chlorides, and density.
@@
 
-# Regression Methods
+# Regression
 ## Multiple Linear Regression
@@
 $$
-Although it is possible to use every predictor in the MLR model, we instead use variable selection to reduce the number of predictors to six (`volatile.acidity`, `chlorides`, `pH`, `total.sulfur.dioxide`, `sulphates`, and `alcohol`) while maintaining predictive accuracy (see Variable Selection section for details).
-Applying the MLR with the reduced model, we get the following:
+Although it is possible to use every predictor in the MLR model, we used best subsets selection to select the variables that best explain the variation in the response.
+Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors.
+Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
+We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
+
+```{r reg.select}
+# Create model and get summary output
+lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
+smre <- lm.sub %>% summary()
+
+# Plot the Cp, BIC, R^2, RSS results in one plot
+par(mfrow = c(2,2))
+plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
+points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
+plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
+points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
+plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
+points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
+plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
+points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
+
+# Best subset selection
+best.betas <- coef(lm.sub, id = 6) %>% round(2)
+```
+
+Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
+Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
+This was the model favored by BIC, which gives the greatest penalty for complexity in models.
+The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
+All of the variables from the classification selection appear here with the exception of density.
+In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses.
+The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.
 
@@
 #### REGRESSION ####
-lm.reg <- lm(quality ~ volatile.acidity + chlorides + pH + total.sulfur.dioxide + sulphates + alcohol, data = trn)
+lm.reg <- lm(quality ~ volatile.acidity + chlorides + 
+               pH + total.sulfur.dioxide + sulphates +
+               alcohol, data = trn)
 summ(lm.reg, digits = 3)
@@
 When residuals are standardized, as in the top right plot, we expect about 95% to be inside the [-1.96, 1.96] interval; the actual proportion is `r round(scale.res.2std, 3)`.
-The Q-Q plot at bottom left compares the normality of residuals against a theoretical distribution.
+The QQ-plot at bottom left compares the normality of residuals against a theoretical distribution.
 Except for a slight downward bias of the lowest 5% of residuals at 2 standard deviations below 0, the data fits a normal distribution almost perfectly.
 On the bottom right plot, high leverage points (those with leverage $\ge$ 0.03) are not dramatically biased, being spread both above and below the 0 residual.
-Cook's Distance top out at about 0.05, well below the generally accepted threshold of 1 needed to quality as highly influential.
+Cook's Distance tops out at about 0.05, well below the generally accepted threshold of 1 that is considered highly influential.
 These plots taken together suggest that the data generally fulfills the requirements for modeling using MLR.
@@
 The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
+Based on the residual plots, though, we are confident that our model does not violate any of the model assumptions.
 
@@
 As mentioned earlier in the classification section, Random Forest is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees.
-This model will be built exactly in the same way as the classification model, except the tree will not be predicting if the wine is a part of the good or bad class, but rather predicts the quality score associated with each wine.
-As before the model will first be built using cross validation to determine the number of variables that should be used in order to produce the smallest test MSE.
-Similar to the random forest classification tree, the cross validation selects the tree with 2 variables.
+This model will be built exactly in the same way as the classification model, except the tree will not be predicting if the wine is a part of the "high" or "low" class, but rather predicts the quality score associated with each wine.
+The predicted value of the quality score will be an average of all of the generated regression trees.
+As before, the model will first be built using cross validation to determine the number of variables that should be used in order to produce the smallest test RMSE.
 
@@
 ```
-The test MSE for the Regression Random Forest using 2 variables is 0.55. COMPARE THIS TO MLR WHEN WE GET YUN'S CODE.
 
+Based on cross validation, we chose our random forest model for regression to have two randomly-selected variables considered at each split. 
 # Results
@@
 Based on the output, the best model for both classification and regression was the random forest.
-This was not a surprise, since the ensemble methods tend to outperform regular methods.
+This supports our hypothesis that the random forest would outperform the other methods.
+This was not necessarily a surprise, since the ensemble methods tend to outperform regular methods.
 As an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
 We suspect that other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracy.
+
+In the regression setting, the values of root mean squared error (RMSE) for both models were relatively low.
+The RMSE for the random forest was `r df.reg[2,2]` and the RMSE for the multiple linear regression model was `r df.reg[1,2]`, further supporting our hypothesis that the random forest would outperfrom the other models.
 
@@
 Based on the results of our study, we were able to predict the high or low quality of a given wine with about 70-80% accuracy.
+Despite not being able to achieve 80% prediction accuracy with the basic classification tree, we *were* able to with the random forest.
 Although this is a *good* result considering the non-information rate of 53%, it is by no means excellent.
 These results indicate that machine learning algorithms could be a good technique when trying to predict wine quality, but that there is room for improvement.
-With a more formal, industry expert-guided variable selection process, we believe that higher prediction accuracies can be achieved.
+With a more formal, industry-expert-guided variable selection process, we believe that higher prediction accuracies can be achieved.
 Machine learning models could potentially be combined with other techniques to add value to winemakers.
@@
 It would be interesting to see how different weather conditions, terroirs, plant ages, and slopes affect the outcome of the wine in terms of quality.
-
-# NOT NEEDED ???
-
-I'm not sure we need these next 2 sections because most of the methods we are using do model selection.
-I don't have Yun's code, but I'm assuming he does some sort of variable selection, so this part seems extraneous.
-
-## Cross Validation
-Again, our focus is on the classification accuracy of machine learning methods.
-In particular, we will be using traditional cross-validation methods to assess the accuracy, sensitivity, and specificity of each of the models.
-We split the data into a training and test data set in order to later evaluate the model's prediction accuracy or root mean squared error (RMSE).
-For this project we used a 80/20 split, training the data on the 80% and testing the trained models on the withheld 20%.
-We will then repeat this process over 5 folds of the data, averaging the results.
-
-```{r CV, eval = F}
-# Specify the cross-validation method (5-fold CV)
-trctrl <- trainControl(method = "cv", number = 5)
-```
-
-## Variable Selection
-Since we have two different tasks that we are trying to complete (classification and regression), we went about selecting the variables for each task differently.
-For the classification task, we ran an initial random forest model on the entire dataset with selected predictors.
-We then looked at the variable importance of the predictors to decide which variables we would use in our final models.
-Within the context of machine learning, variable importance refers to how much a given model "uses" that variable to make accurate predictions.
-In other words, the more a model relies on a variable to make predictions, the more important it is for the model.
-
-```{r class.select, out.height='4in', eval = F}
-# Write preliminary model  
-model.class.pre <- class ~ .
-
-# Fit preliminary random forest on entire data set
-rf.class.pre <- randomForest(model.class.pre, data = wine[,-12], 
-                             ntree = 500, mtry = 3, importance = TRUE)
-
-# Create data frame of variable importance
-df.class.pre <- rf.class.pre$importance %>% as.data.frame()
-X <- rownames(df.class.pre)
-
-# Sort the data frame and prep for plot
-VIrf.pre.class.sort <- df.class.pre[order(df.class.pre[,"MeanDecreaseAccuracy"]), , drop = FALSE]
-VIrf.pre.class.sort$X <- rownames(VIrf.pre.class.sort)
-VIrf.pre.class.sort$X <- factor(VIrf.pre.class.sort$X, levels = VIrf.pre.class.sort$X)
-
-# Plot variable importance
-vimp.class <- ggplot(VIrf.pre.class.sort, aes(x = X, y = MeanDecreaseAccuracy)) +
-                geom_bar(stat = 'identity', position = 'dodge', fill="lightblue") +
-                theme(axis.text.x = element_text(angle = 30, hjust = 1)) +
-                ylab("Mean Decrease in Accuracy") + xlab("Predictor") +
-                ggtitle("Variable Importance")
-vimp.class
-```
-
-We decided to choose the five variables with the most variable importance, leaving us with `alcohol`, `sulphates`, `volatile.acidity`, `density`, and `total.sulfur.dioxide`.
-The fact that density proved to be an important variable for classification was relatively surprising.
-One would think that the wine quality would be determined based on its attributes that are detectable by the human senses.
-For this same reason, alcohol, acidity, and sulphates made sense in terms of variable importance.
-Sulphates are essential for the preservation of the wine, an overly- or underly-acidic wine may be considered poor, and the alcohol content has a connection to the fermentation process, where the must (aka the grape juice before it undergoes fermentation) actually becomes wine.
-
-When looking at the problem in terms of a regression context, we used best subsets selection for feature selection.
-Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors.
-Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
-We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
-
-```{r reg.select, out.height='3.5in', eval = T}
-# Create model and get summary output
-lm.sub <- regsubsets(quality ~ ., data = wine[,-13], nvmax = 11)
-smre <- lm.sub %>% summary()
-
-# Plot the Cp, BIC, R^2, RSS results in one plot
-par(mfrow = c(2,2))
-plot(smre$cp, xlab = "# of Preds", ylab = "Cp")
-points(which.min(smre$cp), smre$cp[which.min(smre$cp)], pch = 8, col = "#CC0000")
-plot(smre$bic, xlab = "# of Preds", ylab = "BIC")
-points(which.min(smre$bic), smre$bic[which.min(smre$bic)], pch = 8, col = "#CC0000")
-plot(smre$adjr2, xlab = "# of Preds", ylab = "Adjusted R^2")
-points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
-plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
-points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
-
-# Best subset selection
-best.betas <- coef(lm.sub, id = 6) %>% round(2)
-```
-
-Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
-Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
-This was the model favored by BIC, which gives the greatest penalty for complexity in models.
-The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
-All of the variables from the classification selection appear here with the exception of density.
-In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses.
-The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.
-
-```{r form, eval = F}
-# Write formula for regression models
-form.reg <- quality ~ volatile.acidity + chlorides + pH + 
-            total.sulfur.dioxide + sulphates + alcohol
-
-# Write formula for classification models
-form.class <- class ~ alcohol + sulphates + density +
-              volatile.acidity + total.sulfur.dioxide
-```

### 2021-04-21 DRAFT.Rmd
Jackie Edits

@@
 
-There has been a push towards the implementation of different analytics solutions within the agriculture industry, especially in the realms of crop health and yield.
+There has been a push towards the implementation of different analytic solutions within the agriculture industry, especially in the realms of crop health and yield.
 These sort of analyses can help farmers reduce waste and improve profits.
@@
 Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality.
+The main question the study looks to answer is can the quality of wine be predicted using a statistical model instead of taste testers.
 
-In this study, the objective is to model and predict the quality of wine using a variety of statistical learning techniques.
+The main objective is to model and predict the quality of wine using a variety of statistical learning techniques.
 This will be done by performing both regression and classification methods, with regression ranking the quality of the wines on a scale of zero to ten, with zero being the lowest quality and ten being the highest quality.
@@
 The research question of interest for this study is whether or not machine learning methods can accurately predict the quality of a wine, independent of using taste testers.
-In the classification setting, we hypothesize that the machine learning methods can correctly classify wines as "high" or "low" quality with 80% accuracy.
-In the context of regression, we hypothesize that the regression techniques used will be able to achieve a low root mean squared error.
-In both settings, though, we expect the random forest to outperform the other methods.
+In the classification setting, the team hypothesizes that the machine learning methods can correctly classify wines as "high" or "low" quality with 80% accuracy.
+In the context of regression, the team hypothesizes that the regression techniques used will be able to achieve a low root mean squared error.
+In both settings, though, the team expects the random forest to outperform the other methods.
 
@@
 # Data
-The data set used is the Wine Quality dataset from the UCI Machine Learning Repository.
+The data set used is the Wine Quality data set from the UCI Machine Learning Repository.
 The data set is comprised of 1600 observations of different variants of the Portuguese "Vinho Verde" red wine.
-For each wine, various physical and chemical features of each wine was measured, including a measurement of quality given by an expert wine taster.
-The objective is to use the different features to predict the quality of the wine using different machine learning techniques.
+For each wine, various physical and chemical features of each wine were measured, including a measurement of quality given by an expert wine taster.
+The objective is to use the different features to predict the quality of the wine using multiple machine learning techniques.
 
 ## Variable Descriptions  
-Each entry in the dataset represents the different metrics of the following attributes of a single type of wine.
+Each entry in the data set represents the different metrics of the following attributes of a single type of wine.
 
@@
 ## Data Exploration
-The first thing that was done to the data upon reading it was to look at histograms of all of the variables. In doing so, we cann get a feel for the distributions of the individual variables.
+After reading the data, the first step was to create and view histograms of all of the variables. In doing so, one can get a feel for the distributions of the individual variables.
 
@@
 
-Most wines are moderate wines, with less excellent amd poor wines.
-The density and pH of the wine seem to be normally distributed, while the sulfur, acid, and sulphate content seem skewed to the right.
-The density varies very little; looking at the scale on the distribution, we can see that it is approximately normally distributed around .997 with a standard deviation of about .002.
+Most wines are classified as moderate wines, with less wines falling in the excellent and poor categories.
+The density and pH of the wine seem to be normally distributed, while the sulfur, acid, and sulphate content appear skewed to the right.
+The density varies very little; looking at the scale on the distribution, one can see that it is approximately normally distributed around 0.997 with a standard deviation of about 0.002.
 The distribution of alcohol content came as no surprise; an average bottle of wine will range from about 10-12% alcohol by volume, with some exceptions.
@@
 
-Next, we went ahead and looked at pairwise scatterplots of the variables.
-This way, one can obtain a greater understanding of how the variables relate to one another, as well as if there is any correlation structure and what it looks like.
+Next, pairwise scatterplots of the variables were created and evaluated.
+This method allows one to obtain a greater understanding of how the variables relate to one another, as well as if there is any correlation structure.
 If the variables are related, there may be problems moving forward, so it is important to know these relationships before the analysis is performed.
@@
 Based on the pairwise scatterplots, one may conclude that pH may be linearly correlated with density, sulphates, and alcohol.
-This leads us to believe that there is some relationship between the chemical properties of the wines.
-here also appears to be a relationship between density and alcohol.
+This shows that there may be some relationship between the chemical properties of the wines.
+There also appears to be a relationship between density and alcohol.
 The following correlation plot gives a numeric summary of the linear relationship between the variables.
@@
 The only linear correlations that are notably high are between alcohol and density.
-This makes sense because the density of alcohol is slightly less than water, meaning that as the concentration of alcohol gets higher, the density will naturally decrease.
+This trend is logical because the density of alcohol is slightly less than water, meaning that as the concentration of alcohol gets higher, the density will naturally decrease.
 Similarly, the sulphates and chlorides seem to have a slightly positive linear relationship, while pH and density have a slightly negative linear relationship.
@@
 The two groups were similar in most of the variables.
-Two notable differences are that the wines classified as high quality had a lower mean value than low quality wines for total sulfur dioxide.
-High quality wines had a mean of 39, which was significantly lower than the mean of the low quality wines at 54.6.
-Another difference to note is the difference in the mean for alcohol.
+A notable difference is that the wines classified as high quality had a lower mean value, 39, compared to that of low quality wines, 54.6, for total sulfur dioxide.
+Another dissimilarity is the difference in the mean for alcohol.
 The high quality wines had a mean alcohol content at 10.8, while the low quality wines had a lower mean alcohol content around 9.9.
-There were also differences in fixed acidity, with the lower quality wines having a lower mean.
-There were also several variables that don't seem to change between the two groups including pH, residual sugar, chlorides, and density.
+Differences also existed in fixed acidity, with the lower quality wines having a lower mean.
+There were also several variables that do not seem to change between the two groups including pH, residual sugar, chlorides, and density.
 
@@
 
-Fortunately, the dataset was very tidy; there were no missing values.
+Fortunately, the data set was very tidy; there were no missing values.
 The data was first split into a training and test data set.
 The training data set uses 80% of the observations and the test set contains the remaining 20%.
-The models are built on the training set and evaluated on the test set.
+The models were built on the training set and evaluated on the test set.
 This is important so that the models are not overfit.
-A variable was then created for classification, transforming the quality variable into a binary "low" or "high" response as mentioned in the introduction.
+A variable was then created for classification, transforming the quality variable into a binary "low" or "high" response as mentioned in the Introduction.
 
@@
 This is done to ensure the data is not being overfit.
-The tradeoff here is that pruning increases the bias.
+The trade-off here is that pruning increases the bias.
 
@@
 A tree is built many times with a different number of splits and the tree that produces the smallest error rate is selected.
-Cross validation on this model shows that 8 is the optimal number of splits to use.
+Cross validation on this model shows that 8 is the optimal number of splits to use, as compared to the first tree which produced 10 splits.
 The resulting pruned model will now have 8 terminal nodes and uses four variables: Alcohol, Total Sulfur Dioxide, Sulphates, and Volatile Acidity.
@@
 # prune.tree.class.test_err
+tree.res
 ```
 
-The pruned tree produces a test error rate to the original tree with a misclassification error of `r prune.tree.class.test_err`.
+The pruned tree produces a test error rate smaller than the original tree with a misclassification error of `r prune.tree.class.test_err`.
 The advantage of using a basic tree is that it is easy to understand and has a good interpretability, which is not something that translates over to ensemble methods.
@@
 That being said, using ensemble methods like the random forest can often improve the error rate of a simple classification tree, since it is an average of many classification trees.
-While the test MSE was small in the classification tree, we decided to investigate the performance of a random forest, hoping to improve the accuracy of our predictions.
+While the test MSE was small in the classification tree, a random forest model was built next, in hopes that it may improve the accuracy of the predictions.
 
@@
 The test MSE is evaluated for each and the smallest value is selected.
-A plot of mtry as a function of the classification error is included below.
+A plot of mtry as a function of the misclassification error is included below.
 
@@
 
-From cross validation, the model that produced the smallest error rate was the model that was built with   `r which.min(test.err)` randomly-chosen variables at each split.
+From cross validation, the model that produced the smallest error rate was the model built with   `r which.min(test.err)` randomly-chosen variables at each split.
 This model produced a test error of `r round(rf.class.test_err, 3)`, which is significantly lower than the test error for the basic classification tree (`r round(prune.tree.class.test_err, 3)`).
-The random forest was selected as the best method for predicting if a wine were to fall into the "high" or "low" quality desgination.
+The random forest was selected as the best method for predicting if a wine were to fall into the "high" or "low" quality designation.
 
@@
 $$
-Although it is possible to use every predictor in the MLR model, we used best subsets selection to select the variables that best explain the variation in the response.
-Using the `leaps` library and the `regsubsets()` function in R, we are able to compare all of the possible models from the given set of predictors.
-Through best subset selection, we can get the set of predictors for a model with any given amount of predictors.
-We looked at three different model criteria to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
+Although it is possible to use every predictor in the MLR model, Best Subsets Selection selected the variables that best explain the variation in the response.
+The `leaps` library and the `regsubsets()` function in R produces comparisons of all possible models from the given set of predictors.
+Best Subset Selection compares all the possible combinations of predictors to select the best model that produces the smallest test error.
+Three different model criteria were evaluated to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
 
@@
 Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
-Since they were so close, we adhered to the Occam's Razor and went with the simplest, six-variable model.
+All three produced a similar error rate, so the theory of Occam's Razor was implemented and the simplest, six-variable model was selected.
 This was the model favored by BIC, which gives the greatest penalty for complexity in models.
 The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
-All of the variables from the classification selection appear here with the exception of density.
-In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, since it is similar to density in the sense that it is not easily detected by the five human senses.
+All of the variables from the classification selection appear with the exception of density.
+In the regression context, `pH` was the variable that proved to be a rather unintuitive inclusion, because it is similar to density in that it is not easily detected by human senses.
 The inclusion of these two variables in the model portray the importance of the chemical properties of a bottle of wine when it comes to quality.
@@
   * Linear relationship between response and predictors
-  * No collinearity between predictor variables
-  * $Y_i$'s are iid
+  * No collinearity exists between predictor variables
+  * $Y_i$'s are independent and identically distributed
   * $\epsilon_i \sim N(0, \sigma^2)$
 
-To assess these assumptions, we looked at various diagnostics plots:
+To assess these assumptions, various diagnostics plots were evaluated:
 ```{r lm.diag}
@@
 
-Residuals indicate that our MLR assumptions are mostly satisfied.
+Residuals indicate that the MLR assumptions are mostly satisfied.
 The residuals vs.
-fitted values plot (top left) has unusual stratified lines, but this is caused by the actual responses being integer values of 4, 5, 6, 7, etc.
+fitted values plot (top left) has unusual stratified lines, but this is caused by the observed responses being integer values of 4, 5, 6, 7, etc.
 rather than continuous values.
-The plot does not show heteroscedasticity, and residuals are spread evenly around 0 over the entire range of fitted values.
-When residuals are standardized, as in the top right plot, we expect about 95% to be inside the [-1.96, 1.96] interval; the actual proportion is `r round(scale.res.2std, 3)`.
-The QQ-plot at bottom left compares the normality of residuals against a theoretical distribution.
+The plot does not show heteroscedasticity and residuals are spread evenly around 0 over the entire range of fitted values.
+When residuals are standardized, as in the top right plot, one would expect about 95% to be inside the [-1.96, 1.96] interval; the actual proportion is `r round(scale.res.2std, 3)`.
+The QQ-plot in the bottom left compares the normality of residuals against a theoretical distribution.
 Except for a slight downward bias of the lowest 5% of residuals at 2 standard deviations below 0, the data fits a normal distribution almost perfectly.
@@
 
-The biggest benefit to using a multiple linear regression model is its interpretability and ease of use.
-The coefficients can be computed efficiently and they are have clear and easy interpretations.
+The largest benefit to using a multiple linear regression model is its interpretability and ease of use.
+The coefficients can be computed efficiently and they have clear and easy interpretations.
 It is also possible to make inference on any of the variables or linear combination of variables.
-The most signficiant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, our estimates and inference will be unstable at best and invalid at worst.
-Based on the residual plots, though, we are confident that our model does not violate any of the model assumptions.
+The most significant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, the estimates and inference will be unstable at best and invalid at worst.
+Based on the residual plots, though, it appears that the model does not violate any of the model assumptions.
 
 ## Random Forest for Regression
-As mentioned earlier in the classification section, Random Forest is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees.
+As mentioned earlier in the Classification section, Random Forest is an ensemble tree-based method, which creates multiple trees from bootstrap samples and averages the results from all the fitted trees.
 This model will be built exactly in the same way as the classification model, except the tree will not be predicting if the wine is a part of the "high" or "low" class, but rather predicts the quality score associated with each wine.
@@
 
-Based on cross validation, we chose our random forest model for regression to have two randomly-selected variables considered at each split. 
+Based on cross validation, the random forest model for regression using two randomly-selected variables considered at each split was selected.
+This model produced a test error rate of `rf.reg.test.err`.
+
 # Results
@@
 Based on the output, the best model for both classification and regression was the random forest.
-This supports our hypothesis that the random forest would outperform the other methods.
+This supports the hypothesis that the random forest would outperform the other methods.
 This was not necessarily a surprise, since the ensemble methods tend to outperform regular methods.
 As an average of many classification trees, ensemble methods generally improve prediction accuracy and reduce overall variance.
-We suspect that other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracy.
+Other ensemble methods, such as bagging or boosting, may also be able to yield higher prediction accuracy.
 
 In the regression setting, the values of root mean squared error (RMSE) for both models were relatively low.
-The RMSE for the random forest was `r df.reg[2,2]` and the RMSE for the multiple linear regression model was `r df.reg[1,2]`, further supporting our hypothesis that the random forest would outperfrom the other models.
+The RMSE for the random forest was `r df.reg[2,2]` and the RMSE for the multiple linear regression model was `r df.reg[1,2]`, further supporting the hypothesis that the random forest would outperform the other models.
 
@@
 The basic classification tree and multiple linear regression model took significantly less time to run than the random forest, but they were less accurate in their predictions.
-If we wanted to run an ensemble method on a very large data set with a lot of features, it would be a computational nightmare.
+If an ensemble method were run on a very large data set with a lot of features, the model would take much longer to run and would be incredibly difficult to interpret.
 
@@
 In particular, the viticulture industry stands to gain a lot from such methods, since historically it has not interacted much with the world of analytics.
-In this study, we went through various machine learning techniques, assessing prediction accuracy in both classification and regression contexts.
+In this study, various machine learning techniques were performed that assessed the prediction accuracy in both classification and regression contexts.
 
-Based on the results of our study, we were able to predict the high or low quality of a given wine with about 70-80% accuracy.
-Despite not being able to achieve 80% prediction accuracy with the basic classification tree, we *were* able to with the random forest.
+Based on the results of the study, prediction of if a given wine has high or low quality was achieved with about 70-80% accuracy.
+Despite not being able to achieve 80% prediction accuracy with the basic classification tree, the random forest model showed an improvement and predicted with a higher accuracy.
 Although this is a *good* result considering the non-information rate of 53%, it is by no means excellent.
 These results indicate that machine learning algorithms could be a good technique when trying to predict wine quality, but that there is room for improvement.
-With a more formal, industry-expert-guided variable selection process, we believe that higher prediction accuracies can be achieved.
+With a more formal, industry-expert-guided variable selection process, the team believes that higher prediction accuracy can be achieved.
 Machine learning models could potentially be combined with other techniques to add value to winemakers.
@@
 The multiple linear regression model is subject to certain distributional and model assumptions, which were not explicitly tested or checked.
-Since the objective of this project was to predict rather than make inference, we didn't explicitly check the model assumptions.
+Since the objective of this project was to predict rather than make inference, model assumptions were not explicitly checked.
 
-All of the wines in the dataset come from wineries in Portugal, and although they are from different wineries, we worry that the results may not be generalizable to wine production globally.
+All wines in the data set came from wineries in Portugal, although they are from different wineries, one may worry that the results may not be generalizable to wine production globally.
 Similarly, will these prediction results hold when it comes to white wine, rose, or sparkling wine?
-Given this variability over location and type of wine, what can we say about the winemaking industry as a whole?
+Given this variability over location and type of wine, what can be said about the wine-making industry as a whole?
 
@@
 For example, regions like Bordeaux and Champagne in France have some of the highest standards in the world for their wines, which are much different than somewhere like Greece or New Zealand.
-How accurate will are predictions be when there are different standards and how might we account for that variation?
+How accurate will the predictions be when there are different standards and how might one account for that variation?
 
 ## Further Research  
-Since this study solely looks at the prediction accuracy of machine learning techniques for red wines, the natural extension of the analysis would to verify that the results hold constant for white wines as well.
+Since this study solely looks at the prediction accuracy of machine learning techniques for red wines, the natural extension of the analysis would be to verify that the results hold constant for white wines as well.
 The University of California, Irvine Machine Learning database has data on white wines, so that would be a good place to start.

### 2021-04-23 DRAFT.Rmd
made the edits suggested by Dr. Griffith: Messed with the histograms and remade them in ggplot Added a references section and citations changed the part on page 4 where we talked about moderate, poor, and excellent classes. made the table rounded to 2 digits instead of a ton

@@
 
-There has been a push towards the implementation of different analytic solutions within the agriculture industry, especially in the realms of crop health and yield.
-These sort of analyses can help farmers reduce waste and improve profits.
+There has been a push towards the implementation of different analytic solutions within the agriculture industry, especially in the realms of crop health and yield (Byrum et al.
+2016).
+These sort of analyses can help farmers reduce waste and improve profits (Dimitriadis 2008).
 In particular, the viticulture industry stands to benefit from such methods, being that the production of wine is lengthy and multifaceted.
-Winemakers can work to optimize revenue by analyzing their input and output (crops and wines).
+Winemakers can work to optimize revenue by analyzing their input and output (crops and wines) (Santesteban 2019).
 Specifically, costs may be cut out if machine learning methods were able to predict the quality of wine based on different metrics, since wine producers would no longer have to hire expert wine tasters to determine wine quality.
@@
 # Create histograms for all 12 variables
-par(mfrow = c(3,4))
-for (i in 1:ncol(wine)){
-  hist(wine[,i], xlab = names(wine)[i], 
-       main = names(wine)[i])
-}
+hist1 <- ggplot(data = wine, aes(x = fixed.acidity)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = 1) +
+           labs(x = "Fixed Acidity", y = "Frequency")
+hist2 <- ggplot(data = wine, aes(x = volatile.acidity)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .1) +
+           labs(x = "Volatile Acidity", y = NULL)
+hist3 <- ggplot(data = wine, aes(x = citric.acid)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .075) +
+           labs(x = "Citric Acid", y = NULL)
+hist4 <- ggplot(data = wine, aes(x = residual.sugar)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = 1) +
+           labs(x = "Residual Sugar", y = NULL)
+hist5 <- ggplot(data = wine, aes(x = chlorides)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .05) +
+           labs(x = "Chlorides", y = "Frequency")
+hist6 <- ggplot(data = wine, aes(x = free.sulfur.dioxide)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = 5) +
+           labs(x = "Free Sulfur Dioxide", y = NULL)
+hist7 <- ggplot(data = wine, aes(x = total.sulfur.dioxide)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = 20) +
+           labs(x = "Total Sulfur Dioxide", y = NULL)
+hist8 <- ggplot(data = wine, aes(x = density)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .001) +
+           labs(x = "Density", y = NULL)
+hist9 <- ggplot(data = wine, aes(x = pH)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .1) +
+           labs(x = "pH", y = "Frequency")
+hist10 <- ggplot(data = wine, aes(x = sulphates)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .1) +
+           labs(x = "Sulphates", y = NULL)
+hist11 <- ggplot(data = wine, aes(x = alcohol)) + 
+           geom_histogram(color = 'black', fill = 'grey', binwidth = .5) +
+           labs(x = "Alcohol", y = NULL)
+hist12 <- ggplot(data = wine, aes(x = factor(quality))) + 
+           geom_bar(color = 'black', fill = 'grey') +
+           labs(x = "Quality", y = NULL)
+grid.arrange(hist1, hist2, hist3, hist4,
+             hist5, hist6, hist7, hist8,
+             hist9, hist10, hist11, hist12,
+             nrow = 3, ncol = 4)
 ```
 
-Most wines are classified as moderate wines, with less wines falling in the excellent and poor categories.
+Most wines in the data set seem to be around the 5 to 7 range in terms of quality rating, meaning that there are not very many excellent or very poor wines.
 The density and pH of the wine seem to be normally distributed, while the sulfur, acid, and sulphate content appear skewed to the right.
@@
 
-kable(good.out, caption = "Summary Output for Good Wines") %>%
+kable(round(good.out, 2), caption = "Summary Output for Good Wines") %>%
   kable_styling(latex_options = "HOLD_position")
-kable(bad.out, caption = "Summary Output for Bad Wines") %>%
+kable(round(bad.out, 2), caption = "Summary Output for Bad Wines") %>%
   kable_styling(latex_options = "HOLD_position")
@@
 # prune.tree.class.test_err
-tree.res
+# tree.res
 ```
@@
 It would be interesting to see how different weather conditions, terroirs, plant ages, and slopes affect the outcome of the wine in terms of quality.
+
+# References
+
+  * Byrum, Joseph, et al. "Advanced analytics for agricultural product development." Interfaces 46.1 (2016): 5-17.
+  * Dimitriadis, Savvas, and Christos Goumopoulos. "Applying machine learning to extract new knowledge in precision agriculture applications." 2008 Panhellenic Conference on Informatics. IEEE, 2008.
+  * Liakos, Konstantinos G., et al. "Machine learning in agriculture: A review." Sensors 18.8 (2018): 2674.
+  * Santesteban, Luis G. "Precision viticulture and advanced analytics. A short review." Food chemistry 279 (2019): 58-62.

### 2021-04-24 DRAFT.Rmd
Added our names to the front page. Fixed a couple small bugs and made minor wording revisions.

@@
 title: "ST563 Final Project"
-author: "Final Project Group 6"
+author:
+  - Final Project Group 6
+  - Mana Azizsoltani, Rong Huang, Yun Ji, Jackie Steffan
 date: "`r format(Sys.time(), '%d %B, %Y')`"
@@
 The basic classification tree is based on partitioning the data into subgroups using simple binary splitting.
-Initially, all objects are considered a single group.
+Initially, all observations are considered a single group.
 Then, the group is repeatedly split into two subgroups based on the criteria of a certain variable.
@@
 ## Multiple Linear Regression
-Multiple Linear Regression is the extension of simple linear regression to a set of multiple predictor variables.
+Multiple linear regression is the extension of simple linear regression to a set of multiple predictor variables.
 The goal of multiple linear regression is to model the linear relationship between a continuous response and two or more predictor variables.
@@
 $$
-Although it is possible to use every predictor in the MLR model, Best Subsets Selection selected the variables that best explain the variation in the response.
+Although it is possible to use every predictor in the MLR model, the method of Best Subsets Selection can help choose the predictors that best explain the variation in the response, for any given model containing a fixed number of predictors.
 The `leaps` library and the `regsubsets()` function in R produces comparisons of all possible models from the given set of predictors.
 Best Subset Selection compares all the possible combinations of predictors to select the best model that produces the smallest test error.
-Three different model criteria were evaluated to compare models of different amounts of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
+Three different model criteria were evaluated to compare models of different numbers of predictor variables: Mallow's Cp, BIC, and $R^2_{\text{adj}}$.
 
@@
 Although each of the three model criteria suggested a different number of predictor variables, they were all within the same region (between 6 and 8 predictors).
-All three produced a similar error rate, so the theory of Occam's Razor was implemented and the simplest, six-variable model was selected.
+All three produced a similar error rate, so Occam's Razor was implemented and the simplest, six-variable model was selected.
 This was the model favored by BIC, which gives the greatest penalty for complexity in models.
-The six variables used in the regression context are then `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
+The six variables used in the regression context are `volatile.acidity`, `chlorides`, `total.sulfur.dioxide`, `pH`, `sulphates`, and `alcohol`.
 All of the variables from the classification selection appear with the exception of density.
@@
 lm.reg.trn.pred <- predict(lm.reg, newdata = trn)
-lm.reg.trn.rmse <- sqrt(mean((lm.reg.trn.pred - tst$quality)^2))
+lm.reg.trn.rmse <- sqrt(mean((lm.reg.trn.pred - trn$quality)^2))
 
@@
 The residuals vs.
-fitted values plot (top left) has unusual stratified lines, but this is caused by the observed responses being integer values of 4, 5, 6, 7, etc.
-rather than continuous values.
+fitted values plot (top left) has unusual stratified lines, but this effect is caused by the observed responses being integer values of 4, 5, 6, 7, etc.
+rather than the continuous values typical of a regression.
 The plot does not show heteroscedasticity and residuals are spread evenly around 0 over the entire range of fitted values.
-When residuals are standardized, as in the top right plot, one would expect about 95% to be inside the [-1.96, 1.96] interval; the actual proportion is `r round(scale.res.2std, 3)`.
+When residuals are standardized, as in the top right plot, one would expect about 95% to be inside the [-1.96, 1.96] interval; the actual percentage is `r round(100*scale.res.2std, 1)`%.
 The QQ-plot in the bottom left compares the normality of residuals against a theoretical distribution.
@@
 On the bottom right plot, high leverage points (those with leverage $\ge$ 0.03) are not dramatically biased, being spread both above and below the 0 residual.
-Cook's Distance tops out at about 0.05, well below the generally accepted threshold of 1 that is considered highly influential.
+Cook's Distance tops out at about 0.08, well below the generally accepted threshold of 1 that is considered highly influential.
 These plots taken together suggest that the data generally fulfills the requirements for modeling using MLR.
@@
 It is also possible to make inference on any of the variables or linear combination of variables.
-The most significant drawback of using the multiple linear regression model is that it is subject to distributional assumptions, meaning that if the assumptions are violated, the estimates and inference will be unstable at best and invalid at worst.
-Based on the residual plots, though, it appears that the model does not violate any of the model assumptions.
+The most significant drawback of using the multiple linear regression model is that the model is subject to distributional assumptions, meaning that if the assumptions are violated, the estimates and inference will be unstable at best and invalid at worst.
+But based on the residual plots it appears that the model does not violate any of the model assumptions.
 
@@
 Based on cross validation, the random forest model for regression using two randomly-selected variables considered at each split was selected.
-This model produced a test error rate of `rf.reg.test.err`.
+This model produced a test error rate of `r round(rf.reg.test_err,4)`.
 
@@
 
-In the regression setting, the values of root mean squared error (RMSE) for both models were relatively low.
+In the regression setting, the test root mean squared error (RMSE) for both models were relatively low.
 The RMSE for the random forest was `r df.reg[2,2]` and the RMSE for the multiple linear regression model was `r df.reg[1,2]`, further supporting the hypothesis that the random forest would outperform the other models.
@@
 Despite not being able to achieve 80% prediction accuracy with the basic classification tree, the random forest model showed an improvement and predicted with a higher accuracy.
-Although this is a *good* result considering the non-information rate of 53%, it is by no means excellent.
+Although this is a *good* result considering the no-information rate of 53%, it is by no means excellent.
 These results indicate that machine learning algorithms could be a good technique when trying to predict wine quality, but that there is room for improvement.

### 2021-04-30 DRAFT.Rmd
Corrected a typo.

@@
 points(which.max(smre$adjr2), smre$adjr2[which.max(smre$adjr2)], pch = 8, col = "#CC0000")
-plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
-points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
+#plot(smre$rss, xlab = "# of Preds", ylab = "Training RSS")
+#points(which.min(smre$rss), smre$rss[which.min(smre$rss)], pch = 8, col = "#CC0000")
 
@@
 Based on cross validation, the random forest model for regression using two randomly-selected variables considered at each split was selected.
-This model produced a test error rate of `r round(rf.reg.test_err,4)`.
+This model produced a test RMSE of `r round(rf.reg.test_err,4)`.
 
